### 2019-05-29 project1/paper.tex
Finishg papers abstract

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}		% Can be removed after putting your text content
\usepackage{textgreek}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{algpascal}
\usepackage{graphicx}
%\graphicspath{ {./images/} }

\title{Replicating Sutton's TD($\lambda$) Algorithm}

\date{June 1st, 2019}	% Here you can change the date presented in the paper title
%\date{} 					% Or removing it

\author{
  Carlos Souza\\
  %Department of Computer Science\\
  Georgia Institute of Technology\\
  São Paulo, SP, Brazil \\
  \texttt{souza@gatech.edu} \\
}

\begin{document}
\maketitle

\begin{abstract}
    TD($\lambda$) is a core algorithm of modern reinforcement learning.
    The purpose of this paper is to replicate results obtained in Richard Sutton's 1988 Learning to Predict by the Methods of Temporal Differences, in which he introduced temporal difference algorithms, theories and insightful experiments.
    We review Sutton's methods, replicating the experiment protocols he used to generate the overall same results he achieved in 1988 in the Random Walk Markov Reward Process.
    Finally, we present and discuss the results obtained, and propose a video animation to facilitate visualizing how differently TD(0) and TD($\lambda$) propagates prediction levels throughout the learning process.
\end{abstract}

% keywords can be removed
\keywords{Reinforcement learning \and Temporal-difference (TD) learning \and Markov reward process}

\section{Introduction}
\label{sec:introduction}
Temporal-difference (TD) learning is a core learning technique in modern reinforcement learning (\cite{Sutton1988}, \cite{Sutton:1998:IRL:551283}, \cite{KLMSurvey}, \cite{Szepesvari:2010:ARL:1855083}).
TD learning generates good estimates for expected returns by quickly bootstrapping from other expected-return estimates.
TD($\lambda$) is one of the most popular algorithms in TD learning.
It was first introduced in Sutton's article \emph{Learning to Predict by the Methods of Temporal Differences}, one of the most referenced articles in the field with over 5,000 citations.
The purpose of this paper is to discuss the results obtained by replicating Sutton's TD($\lambda$) algorithm, providing a complete description of the experiments, implementation details and outcomes.

\subsection{Background}
\label{subsec:background}
Our focus in this paper are \emph{Markov Reward Processes}.
MRPs are stochastic processes that can be described by \(\langle\mathcal{S}, p, r, \gamma\rangle\), where \(\mathcal{S}\) is the set of all possible states, \(p(s'|s)\) is the transition probability from state \(s \in \mathcal{S}\) to state \(s' \in \mathcal{S}\), \(r(s, s')\) is the reward obtained in the transition, and \(\gamma\) is the discount factor.

We will also focus on episodic MRPs, processes that contain \emph{terminal states}.
These states divide the sequences of state transitions into \emph{episodes}.
Once a terminal state is reached, the episode ends and the state is reset to its initial value.
The return at time step \(t\) is the discounted sum of all the rewards observed after this time step until the end of the episode:

\begin{equation}
    G_{t} = \sum_{i=1}^{T-t} \gamma^{i-1} R_{i+1}\label{eqn:1}
\end{equation}

where \emph{T} is the time step of the terminal state.

Our goal is to learn the state-value function $v$ that maps each state $s \in \mathcal{S}$ to the expected return:

\begin{equation}
    v(s) = \mathbb{E}\{G_{t} \mid S_{t} = s\}\label{eqn:2}
\end{equation}

One of the simplest ways to solve a MRP is using Monte Carlo methods.
These methods solve reinforcement learning problems by using experience from interactions with the environment to update their estimate $V$ of $v$ for the nonterminal state $S_{t}$ ocurring in the experience.
Monte Carlo methods wait until the end of the episode, until the full return following the visit is known, to use this value to update current estimate of the state-value function:

\begin{equation}
    V(S_{t}) \leftarrow V(S_{t}) + \alpha(G_{t} - V(S_{t}))\label{eqn:3}
\end{equation}

where $\alpha$ is the learning-rate.
The problem with Monte Carlo methods is that $G_{t}$ is only known by the end of the episode, which means these methods must wait until the end of the episode to determine the increment to $V(S_{t})$.
TD methods, in other hand, need to wait only until the next step $t + 1$.
The simplest TD method, TD(0), uses the observed reward $R_{t+1}$ and the current estimate $V(S_{t+1})$ from the next time step to update value-function estimate of current state:

\begin{equation}
    V(S_{t}) \leftarrow V(S_{t}) + \alpha \underbrace{(R_{t+1} + \gamma V(S_{t+1}) - V(S_{t}))}_{\delta \text{, known as TD error}} \label{eqn:4}
\end{equation}

$TD(0)$, also called \emph{one-step TD}, is just an instance of a more general class of algorithms called TD($\lambda$), where $\lambda = 0$.
The general TD($\lambda$) unify and generalize TD and Monte Carlo methods, producing methods that have Monte Carlo at one extreme ($\lambda = 1$) and one-step TD at the other ($\lambda = 0$).
Its update rule is similar to the TD(0):

\begin{equation}
    V(S_{t}) \leftarrow V(S_{t}) + \alpha (R_{t+1} + \gamma V(S_{t+1}) - V(S_{t})) \: e(S_{t})\label{eqn:5}
\end{equation}

however, it is applied to \emph{all states} accordingly to their \emph{eligibility} $e(S_{t})$.
The eligibility of a state is the degree to which it has been visited in the past.
It is a short-term memory vector that is bumped in the component correspondent to the state recently visited, then fading away in proportion to $\lambda$.
Eligibility can be updated incrementaly as:

\begin{equation}
    e(s) \leftarrow \begin{cases}
                            \gamma \lambda \: e(s) + 1 &\text{if $s =$ current state}\\
                            \gamma \lambda \: e(s) &\text{otherwise}\label{eqn:6}
    \end{cases}
\end{equation}

We can use equations (\ref{eqn:5}) and (\ref{eqn:6}) to devise TD($\lambda$) algorithm:

\alglanguage{pseudocode}
\begin{algorithm}
    \caption{Value-state prediction with TD($\lambda$)}
    \label{alg:1}
    \begin{algorithmic}
        \State Initialize state-value function $V$ arbitrarily
        \For{$episode \gets 1,\, M$}
            \State Initialize $S$
            \State Initialize $e$ to $0$ for all $s \in \mathcal{S}$
            \For{$t \gets 1,\, T$} \Comment{For each time step $t$ in the $episode$}
                \State Observe reward $R_{t+1}$ and next state $S_{t+1}$
                \State $e(S_{t}) \gets e(S_{t}) + 1$
                \State $\delta \gets R_{t+1} + \gamma V(S_{t+1}) - V(S)$
                \State $V(s) \gets V(s) + \alpha \: \delta \: e(s)$ for all $s \in \mathcal{S}$
                \State $e(s) \gets \gamma \: \lambda \: e(s)$ for all $s \in \mathcal{S}$
                \State $S_{t} \gets S_{t+1}$
            \EndFor
        \EndFor
    \end{algorithmic}
\end{algorithm}

\subsection{Random Walk}
\label{subsec:randomwalk}
Sutton (1988) illustrates TD methods with a random-walk example, one of the simplest dynamic systems.
A bounded random walk is a Markov Reward Process.
The episodes always start in the middle state $D$.
At each time step, the walk moves to the neighbor state, either left or right, with equal probability.
The episode ends when walk reaches either extreme states $A$ or $G$.
If termination happens by reaching the extreme right ($G$), environment provides a reward of +1;
all the other rewards are zero.

\begin{figure}[t]
    \includegraphics[scale=0.3]{./images/random_walk.eps}
    \centering
    \caption{A bounded random walk generator used in Sutton's original paper.
    All episodes start in state $D$.
    At each time step, walk has equal probability of either moving left or right.
    Episode ends when walk reaches either $A$ or $G$.
    In case of episode terminating in $G$, +1 reward is received;
    otherwise, reward is zero.}
    \label{fig:1}
\end{figure}

A typical episode terminating in the extreme right might be the following state-reward sequence: $D, 0, C, 0, D, 0, E, 0, F, 0, G, 1$.
An example of a walk ending in the extreme left might be: $D, 0, C, 0, B, 0, A, 0$.
This MRP is undiscounted ($\gamma = 1$);
therefore, the true value of each state is the probability of terminating on the extreme right by starting from that specific state.
Thus, the true values from all states, from $A$ to $G$ respectively, are

\begin{equation}
    v_{true} = \Bigg[0, \frac{1}{6}, \frac{2}{6}, \frac{3}{6}, \frac{4}{6}, \frac{5}{6}, 1\Bigg]
    \centering
\end{equation}

We applied TD($\lambda$) algorithm to effectively predict these true values, replicating Sutton's original paper methods and results.
The following sections explain in details how we implemented the experiments, their results, and compares them with the previously published outcomes.

\section{Methods}
\label{sec:methods}
To replicate Sutton's results, we followed the exact protocols described in the article.
First, we generated the data to be used by all learning procedures, by simulating 100 training sets, each containing 10 episodes (or sequences).
Then, we implemented the experiments using this data.
To measure how good the predictions were, we used the root mean squared (RMS) error between the predictions and the true values.

While implementing the experiments, we changed how the algorithm was written.
Sutton developed all the math in his paper and the algorithms considering that a prediction $P_{t}$ at time step $t$ is function of i) a vector $x_{t}$ of observations available at this time step (representing thus the state), and ii) a vector of modifiable weights, $w$.
This means he expressed all learning procedures and equations as rules for updating $w$.
However, the same learning procedures and equations can also be expressed as rules for updating state-value function $V$ directly, as $w = [V(S_{1}), V(S_{2}), V(S_{3}), ...]$, i.e. the vector in which each element is the state-value function $V(s)$ for each $s \in \mathcal{S}$.
As the underlying math is exactly the same, we arrived at the same results and conclusions, as we will show in the next section.

The first experiment assessed how different values of $\lambda$ produced different predictions with different corresponding RMS errors.
We used $\lambda = 0, 0.1, 0.3, 0.5, 0.7, 0.9$ and $1$.
There are 2 special cases in these values: 0 and 1.
TD(0) is the \emph{one-step} TD, which updates the value-function using the reward from only one step ahead.
TD(1) is an instance of the Monte Carlo methods, which updates value-function only after the episode is complete, using all the rewards until the end of the episode.
This case of $\lambda = 1$ is also referred as Widrow-Hoff in the original paper.

In this experiment, the state-value function was not updated neither online (i.e. after each time step, as shown in the algorithm \ref{alg:1}) nor by the end of each episode.
Instead, the function was only updated after a full loop over all episodes in the training set.
Also, following the paper, we implemented the \emph{repeated presentations} training paradigm in this experiment: we executed the training loop for each training set repeatedly, using the same data, until the state-value function did not change significantly.
Here the paper is unclear about: i) the learning rate ($\alpha$) used, ii) how to assess significant changes in the state-value function throughout the simulation, and iii) the minimum threshold difference between different estimates of state-value functions, below which the simulation ends.
We used $\alpha = 0.01$ (as the paper mentions it used a small learning rate to ensure conversion), RMS error as a way to assess with a single scalar the difference between two estimates of value-functions, and $\theta = 10^{-4}$ as threshold to stop the simulation: whenever the RMS error between two estimates of different value-functions fall below this, the simulation ends.

In the second experiment we simulated different values of $\alpha$ and different values of $\lambda$, to understand the impact of the learning rate in TD.
Three important changes were made in comparison to the first experiment.
First, the training loop was executed only once for each training set.
Second, the state-value function was updated by the end of each episode.
Third, different than the first experiment, where the state-value function was initialized randomly, now the state-value function was initialized at constant value of $0.5$ for all states.
We varied $\alpha$ from $0$ to $0.6$ in steps of $0.05$, and $\gamma$ from $0$
 to $1$ in steps of $0.1$.

\section{Results and Discussion}
\label{sec:results}
\begin{figure}[t]
    \includegraphics[width=\textwidth]{./images/figure.png}
    \centering
    \label{fig:fig2}
    \caption{Replication of Figures 3, 4 and 5 from Sutton's original paper.
    In the left, average error on random walk problem under repeated presentations.
    In the center, average error on random walk problem after 10 episodes for different values of $\alpha$ and $\gamma$.
    In the right, average error at best $\alpha$ on random walk problem for different values of $\gamma$.}
\end{figure}

Figure 2 summarizes the results obtained by replicating Sutton's experiments.
In the left, the chart shows the results of the first experiment, where RMS errors were calculated between the state-value function true values and the ones estimated after repeated presentations until convergence, for different values of $\lambda$.
We averaged the errors over 100 training sets, exactly as done by Sutton.
This chart replicates Figure 3 in the original article.
Exactly as demonstrated by Sutton, we see, after running the first experiment, that lower values of $\lambda$ produced smaller RMS errors.
In fact, the best $\lambda$ was zero, and the worst was 1.
Despite the fact Widrow-Hoff procedure (TD(1)) is an optimal method to minimize error between predictions and actual outcomes in a training set, TD(0) converged to a better state-value function than TD(1).
That's because, when presented repeatedly to a finite training set, TD(0) converges to the \emph{optimal estimates}, while Widrow-Hoff (TD(1)) converges \emph{to the estimates that minimize the error on the training set}, which are different from the optimal estimates.

In the center, the chart shows average errors obtained in a single run of each training set, averaged over 100 training sets, for different values of $\alpha$ and $\lambda$.
This chart replicates Figure 4 in the original article.
Here, again we confirmed Sutton's empirical results.
We showed how TD methods produced smaller errors for all $\lambda < 1$, when compared to Widrow-Hoff (TD(1)).
We also showed how different learning rates produced significantly different state-value function estimates, being the best learning rates around 0.2, as also shown by Sutton.
As expected, we showed how too small learning rates made all learning procedures generate similar errors (i.e. all the different procedures did not learn successfully with smaller values of $\alpha$).
And we could observe as well how too large learning rates caused all learning procedures' errors to explode, confirming our expectation that learning rates cannot be neither too small nor too large for learning algorithms to work.

Finally, in the right, we can see the chart that replicates Figure 5 in Sutton's article.
The results were produced by experiment 2, after a single run of each training set, also averaged over 100 training sets, for different values of $\alpha$ and $\lambda$.
For each value of $\lambda$, we selected the best value of $\alpha$.
Again, our results matched Sutton's: all estimates of state-value function for $\lambda < 1$ produced lower errors than in the $\lambda = 1$ case.
However, as shown by Sutton, the best case was not $\lambda = 0$, but a value for $\lambda$ around 0.3.
That's because TD(0) is slow at propagating prediction levels back throughout the episode, given eligibility factor.
TD algorithm uses the eligibility factor, which in case of $\lambda = 0$, updates only the previously visited state in a given episode.
For $\lambda > 0$, eligibility updates not only the previously visited state, but all states proportionally, which propagates the prediction levels faster.
This is not a problem when the training set is repeatedly presented, but that was not the case in experiment 2, when the training set was seen only once by the learning algorithm.

To try to visualize how fast TD($\lambda$) propagates prediction levels to other states in comparison to TD(0), for any $\lambda > 0$, we created a video animation of experiment 2 protocol, which can be seen in the following Youtube link:
\begin{center}
    \url{https://youtu.be/mBqyQpL8_Vc}
\end{center}

The animation shows experiment 2 protocol, a single run of each training set, for the 100 training sets, for 2 TD learning procedures in parallel.
The first, showed in the first column of charts, is a simulation of TD(0).
The second, in the second column, simulates TD($\lambda = 0.3$).
Both simulations run in parallel, using exactly the same data and same learning rate (0.2).
We used colors to see how each state-value estimate is changed at each time step.
The darker the color, the more change it had versus the previous time step.
This animation, played back in different speeds, helped us visually confirm Sutton's last point about how TD(0) is slow at propagating predictions.
\begin{figure}[t]
    \includegraphics[width=\textwidth]{./images/screenshot.png}
    \centering
    \label{fig:fig3}
    \caption{Screenshot of the video animation showing experiment 2 protocol, performed for 2 TD learning procedures in parallel.
    The first, TD(0), is shown in first column of charts.
    The second, TD($\lambda$) for $\lambda = 0.3$, is shown in the second column of charts.
    In the first row of charts, we show the current estimate of the state-value function for each of the TD learning procedures.
    In the second row, we show the eligibility of each state at that specific time step, for each of the TD learning procedures (TD(0) in blue, TD($\lambda = 0.3$) in red).
    The upper right corner of the screenshot shows specific information about the current state of the simulation.
    The lower right corner of the screenshot shows the RMS error for each of the 10 episodes, averaged over all training sets seen up until that specific moment.}
\end{figure}

\section{Conclusion}
\label{sec:conclusion}
In this work, we implemented TD($\lambda$) reinforcement learning algorithm, replicating Figures 3, 4 and 5 from Sutton's 1998 paper \emph{Learning to Predict by the Methods of Temporal Differences}.
Our results matched the results presented in the paper, and we were able to confirm all main points made in the article about the random walk experiments.

\bibliographystyle{unsrt}
\bibliography{references}  %%% Remove comment to use the external .bib file (using bibtex).
%%% and comment out the ``thebibliography'' section.

\end{document}

### 2019-05-29 project1/final_solution.py

@@ -479,6 +479,7 @@ class Plot(object):
             ax.set_xticks(df.index.values)
             ax.annotate('Widrow-Hoff', xy=(df.iloc[-1].name, df.iloc[-1].values[0]),
                          xytext=(-105, -8), textcoords='offset pixels', fontsize=10)
+            ax.set_facecolor('ivory')
 
         if self.save_to_file:
             df.to_csv('images/figure1.csv')
@@ -522,6 +523,7 @@ class Plot(object):
             labels = [] if legend is None else [str(x._text) for x in legend.texts]
             handles = [] if legend is None else legend.legendHandles
             ax.legend(reversed(handles), reversed(labels), frameon=False)
+            ax.set_facecolor('ivory')
 
         if self.save_to_file:
             df.to_csv('images/figure2.csv')
@@ -561,6 +563,7 @@ class Plot(object):
             ax.set_xlabel(r'$\lambda$', fontsize=18) #
             ax.annotate('Widrow-Hoff', xy=(df.index[-1], df.iloc[-1, -1]),
                          xytext=(-105, -8), textcoords='offset pixels', fontsize=10)
+            ax.set_facecolor('ivory')
 
         if self.save_to_file:
             df.to_csv('images/figure3.csv')

### 2019-05-29 project1/final_solution.py
Adding requirements

@@ -12,11 +12,8 @@ used in the report. It also creates them in individual files,
 replications of Sutton's Figure 3, 4 and 5 respectively.
 
 Todo:
-    * Write report
     * Write README.md
     * Test in ubuntu
-    * Include a requirements.txt in the package
-    * Create folder 'images' if it doesn't exist
 
 Created by Carlos Souza (souza@gatech.edu). May-2019.
 
@@ -32,6 +29,7 @@ import random
 from collections import defaultdict
 from time import sleep
 import argparse
+import os
 
 
 class Environment(object):
@@ -433,15 +431,17 @@ class Plot(object):
         seed (int): If True, sets the seed. Useful to reproduce exact results.
         save_to_file (bool): If True, saves charts in files. Otherwise shows in
             screen.
+        images_folder (str): Folder where to save the images.
 
     Attributes:
         solver (Solver): Solver object used in performing experiments.
         data (list): Dataset with all training sets used in the simulations.
         save_to_file (bool): If True, saves charts in files. Otherwise shows in
             screen.
+        images_folder (str): Folder where to save the images
 
     """
-    def __init__(self, seed=None, save_to_file=True):
+    def __init__(self, seed=None, save_to_file=True, images_folder='images'):
         self.solver = Solver()
         if seed is not None:
             np.random.seed(seed)
@@ -449,6 +449,9 @@ class Plot(object):
 
         self.data = self.solver.generate_data()
         self.save_to_file = save_to_file
+        self.images_folder = images_folder
+        if not os.path.exists(images_folder):
+            os.makedirs(images_folder)
 
     def figure1(self, alpha=0.01, lambdas=[0.0, 0.1, 0.3, 0.5, 0.7, 0.9, 1.0],
                 axis=None):
@@ -482,8 +485,8 @@ class Plot(object):
             ax.set_facecolor('ivory')
 
         if self.save_to_file:
-            df.to_csv('images/figure1.csv')
-            fig.savefig('images/figure1.png')
+            df.to_csv(f'{self.images_folder}/figure1.csv')
+            fig.savefig(f'{self.images_folder}/figure1.png')
         else:
             plt.show()
 
@@ -526,8 +529,8 @@ class Plot(object):
             ax.set_facecolor('ivory')
 
         if self.save_to_file:
-            df.to_csv('images/figure2.csv')
-            fig.savefig('images/figure2.png')
[…]

### 2019-05-29 project1/final_solution.py
Adding README.md and space to put git hash

@@ -12,7 +12,6 @@ used in the report. It also creates them in individual files,
 replications of Sutton's Figure 3, 4 and 5 respectively.
 
 Todo:
-    * Write README.md
     * Test in ubuntu
 
 Created by Carlos Souza (souza@gatech.edu). May-2019.

### 2019-05-30 project1/final_solution.py
Fixing Matplotlib bug to run in Ubuntu

@@ -11,17 +11,15 @@ used in the report. It also creates them in individual files,
 'images/figure1.png', 'images/figure2.png', and 'images/figure3.png', which are
 replications of Sutton's Figure 3, 4 and 5 respectively.
 
-Todo:
-    * Test in ubuntu
-
-Created by Carlos Souza (souza@gatech.edu). May-2019.
+Created by Carlos Souza (souza@gatech.edu)
+May-2019
 
 """
 
 import numpy as np
 import pandas as pd
-#import matplotlib
-#matplotlib.use('Agg')
+import matplotlib
+matplotlib.use('Agg')
 import matplotlib.pyplot as plt
 from tqdm import tqdm
 import random

### 2019-05-30 project1/final_solution.py
Adding watermark plot capability

@@ -11,6 +11,9 @@ used in the report. It also creates them in individual files,
 'images/figure1.png', 'images/figure2.png', and 'images/figure3.png', which are
 replications of Sutton's Figure 3, 4 and 5 respectively.
 
+TODO:
+    * Test again in Linux (after conditional importing)
+
 Created by Carlos Souza (souza@gatech.edu)
 May-2019
 
@@ -18,8 +21,11 @@ May-2019
 
 import numpy as np
 import pandas as pd
-import matplotlib
-matplotlib.use('Agg')
+from sys import platform
+if platform == 'linux':
+    import matplotlib
+    matplotlib.use('Agg')
+
 import matplotlib.pyplot as plt
 from tqdm import tqdm
 import random
@@ -580,7 +586,7 @@ class Plot(object):
         self.solver.experiment_2_live(self.data, file_v, file_e, file_error,
                                       file_annotation, file_evol)
 
-    def generate_all(self):
+    def generate_all(self, watermark=False):
         """Generate all charts
 
         This function runs all experiments and generates all charts used in the
@@ -597,6 +603,10 @@ class Plot(object):
         print('\nGenerating Figure 3...')
         self.figure3(axis=ax3)
         print('\nDONE!')
+        if watermark:
+            font = {'color':  'gray', 'weight': 'normal', 'size': 100, 'alpha': 0.2}
+            fig.text(0.15, 0.5, 'Carlos Souza', fontdict=font)
+
         if self.save_to_file:
             fig.savefig(f'{self.images_folder}/figure.png')
         else:
@@ -609,10 +619,15 @@ if __name__ == '__main__':
                         help='Use this flag to generate DataFrames to create'
                              'video simulation. Check README.md to see how to'
                              'create the video simulation.')
+    parser.add_argument('-s', '--show', action="store_true", default=False,
+                        help='Use this flag to show image on the screen '
+                             'instead of saving it to a file.')
+    parser.add_argument('-w', '--watermark', action="store_true", default=False,
+                        help='Use this flag to create images with watermark.')
     args = parser.parse_args()
 
-    p = Plot(seed=200357)
+    p = Plot(seed=200357, save_to_file=not args.show)
     if not args.live:
-        p.generate_all()
+        p.generate_all(watermark=args.watermark)
     else:
         p.live_plot()

### 2019-06-01 project1/final_solution.py
Adding new random walk illustration and fixing matplotlib bug in linux environment

@@ -12,7 +12,8 @@ used in the report. It also creates them in individual files,
 replications of Sutton's Figure 3, 4 and 5 respectively.
 
 TODO:
-    * Test again in Linux (after conditional importing)
+    * Adjust random_walk illustration and text(from A to E, to be compatible
+      with video screenshot)
 
 Created by Carlos Souza (souza@gatech.edu)
 May-2019

### 2019-06-01 project1/live_plot.py

@@ -1,4 +1,9 @@
 import pandas as pd
+from sys import platform
+if platform == 'linux':
+    import matplotlib
+    matplotlib.use('Agg')
+    
 import matplotlib.pyplot as plt
 import matplotlib.animation as animation
 from pandas.errors import EmptyDataError

### 2019-06-01 project_1/live_plot.py
Reorganizing folders to commit to GT server

@@ -1,9 +1,41 @@
+# -*- coding: utf-8 -*-
+"""OMSCS Reinforcement Learning - CS-7642-O03 - Project #1 Solution
+
+This code solves Project #1 from OMSCS Reinforcement Learning - CS-7642-O03,
+generating the video referenced in report, which can be seen in the following
+Youtube link
+
+    https://youtu.be/mBqyQpL8_Vc
+
+To re-create it, you will have to run 2 commands in parallel: i) one to capture
+the live data being produced by the simulation (and either display it in the
+screen or save it in `images/animation.mp4`), and ii) other to generate the live
+data. To do (i), first run:
+
+    $ python live_plot.py
+
+Then, to generate the live data (ii), then run in parallel:
+
+    $ python final_solution.py --live
+
+The default behaviour of `live_plot.py` program is to save the video in the file
+`images/animation.mp4`. If instead you want to see it in the screen, run
+
+    $ python live_plot --show
+
+instead.
+
+Created by Carlos Souza (souza@gatech.edu)
+May-2019
+
+"""
+
 import pandas as pd
 from sys import platform
 if platform == 'linux':
     import matplotlib
     matplotlib.use('Agg')
-    
+
 import matplotlib.pyplot as plt
 import matplotlib.animation as animation
 from pandas.errors import EmptyDataError

### 2019-06-01 reinforcement-learning-an-introduction/chapter02/ten_armed_testbed.py

@@ -1,232 +1,0 @@
-#######################################################################
-# Copyright (C)                                                       #
-# 2016-2018 Shangtong Zhang(zhangshangtong.cpp@gmail.com)             #
-# 2016 Tian Jun(tianjun.cpp@gmail.com)                                #
-# 2016 Artem Oboturov(oboturov@gmail.com)                             #
-# 2016 Kenta Shimada(hyperkentakun@gmail.com)                         #
-# Permission given to modify the code as long as you keep this        #
-# declaration at the top                                              #
-#######################################################################
-
-import matplotlib
-matplotlib.use('Agg')
-import matplotlib.pyplot as plt
-import numpy as np
-from tqdm import tqdm
-
-class Bandit:
-    # @k_arm: # of arms
-    # @epsilon: probability for exploration in epsilon-greedy algorithm
-    # @initial: initial estimation for each action
-    # @step_size: constant step size for updating estimations
-    # @sample_averages: if True, use sample averages to update estimations instead of constant step size
-    # @UCB_param: if not None, use UCB algorithm to select action
-    # @gradient: if True, use gradient based bandit algorithm
-    # @gradient_baseline: if True, use average reward as baseline for gradient based bandit algorithm
-    def __init__(self, k_arm=10, epsilon=0., initial=0., step_size=0.1, sample_averages=False, UCB_param=None,
-                 gradient=False, gradient_baseline=False, true_reward=0.):
-        self.k = k_arm
-        self.step_size = step_size
-        self.sample_averages = sample_averages
-        self.indices = np.arange(self.k)
-        self.time = 0
-        self.UCB_param = UCB_param
-        self.gradient = gradient
-        self.gradient_baseline = gradient_baseline
-        self.average_reward = 0
-        self.true_reward = true_reward
-        self.epsilon = epsilon
-        self.initial = initial
-
-    def reset(self):
-        # real reward for each action
-        self.q_true = np.random.randn(self.k) + self.true_reward
-
-        # estimation for each action
-        self.q_estimation = np.zeros(self.k) + self.initial
-
-        # # of chosen times for each action
[…]

### 2019-06-01 homeworks/hw1/environment.py
Finish fixing folders to submit into GT server

@@ -1,99 +1,0 @@
-import numpy as np
-from random import randint
-from numpy.random import choice
-import pprint
-
-
-class Environment(object):
-	def __init__(self, mask):
-		self.mask = mask
-		self.bankroll = 0
-		self.count = 0
-		self.num_actions = 2
-
-	def __str__(self):
-		return f'N = {len(self.mask)}, isBadSide = {self.mask}'
-		
-	def step(self, action):
-		"""Step function
-
-		Args:
-			action: The action being taken
-
-		Returns:
-			tuple: Containing next state, reward, done and info
-
-		"""
-		self.count += 1
-		if action == 0:
-			return self.bankroll, 0, True, (self.count, None)
-		
-		roll = randint(1, len(self.mask))
-		if self.mask[roll - 1] == 1:
-			return 0, -self.bankroll, True, (self.count, roll)
-		else:
-			self.bankroll += roll
-			return self.bankroll, roll, False, (self.count, roll)
-		
-	def reset(self):
-		self.bankroll = 0
-		self.count = 0
-		return self.bankroll
-
-	def get_probs(self, state, action):
-		"""Get transition probabilities after taking action in a given state
-
-		Args:
-			state: The current state
-			action: The action taken
-
-		Returns:
-			list of tuples: A list of tuples containing probability, next_state,
-				reward and done
-
-		"""
-		#if self.mask[state - 1] == 1:
-		#	return [(1, 0, -state, True)]
-
-		if action == 0:
-			return [(1, state, 0, True)]
-
-		result = []
-		for roll in range(1, len(self.mask) + 1):
-			if self.mask[roll - 1] == 1:
-				leaf = (1.0 / len(self.mask), 1000, -state, True)
-			else:
-				leaf = (1.0 / len(self.mask), state + roll, roll, False)
-
-			result.append(leaf)
-
-		return result
-
-
-def q_from_v(env, V, s, gamma=1):
-	q = np.zeros(2)
-	for a in range(2):
-		for prob, next_state, reward, done in env.P[s][a]:
-			q[a] += prob * (reward + gamma * V[next_state])
-	return q
-
-		
-if __name__ == '__main__':
-	pp = pprint.PrettyPrinter()
-	env = Environment([1, 1, 1, 0, 0, 0])
-	pp.pprint(env.get_probs(4, 1))
-
-	exit(0)
-
-	for i_episode in range(20):
-		print(f'Episode {i_episode + 1}')
-		state = env.reset()
[…]

### 2019-06-01 homeworks/hw1/final_solution.py

@@ -1,76 +1,0 @@
-import numpy as np
-from collections import defaultdict
-
-
-def get_probs(mask, state, action):
-    """Get transition probabilities after taking action in a given state
-
-    Args:
-        state: The current state
-        action: The action taken
-
-    Returns:
-        list of tuples: A list of tuples containing probability, next_state,
-            reward and done
-
-    """
-    if action == 0:
-        return [(1, state, 0, True)]
-
-    result = []
-    for roll in range(1, len(mask) + 1):
-        if mask[roll - 1] == 1:
-            leaf = (1.0 / len(mask), 1000, -state, True)
-        else:
-            leaf = (1.0 / len(mask), state + roll, roll, False)
-
-        result.append(leaf)
-
-    return result
-
-
-def get_possible_states(mask, origin=[0]):
-    result = []
-    for node in origin:
-        for i in range(1, len(mask) + 1):
-            if (mask[i - 1] == 0) and ((node + i) not in result) and ((node + i) not in origin):
-                result.append(node + i)
-
-    return origin + result
-
-
-def q_from_v(mask, V, s, gamma=1):
-    q = np.zeros(2)
-    for a in range(2):
-        for prob, next_state, reward, done in get_probs(mask, s, a):
-            q[a] += prob * (reward + gamma * V[next_state])
-    return q
-
-
-if __name__ == '__main__':
-    #mask = [1, 1, 1, 0, 0, 0]
-    #mask = [1,1,1,1,1,1,0,1,0,1,1,0,1,0,1,0,0,1,0,0,1,0]
-    #mask = [1,1,1,1,0,0,0,0,1,0,1,0,1,1,0,1,0,0,0,1,0]
-    mask = [0,0,0,0,0,0,0,0,0,1,0,0,1,1,0,0]
-    previous = [0]
-    for i in range(4):
-        states = get_possible_states(mask, origin=previous)
-        previous = states
-
-    states += [1000]
-    print(states)
-
-    gamma = 1
-    theta = 1e-8
-    V = defaultdict(lambda: 0) #np.zeros(len(states))
-    while True:
-        delta = 0
-        for s in states:
-            v = V[s]
-            V[s] = max(q_from_v(mask, V, s, gamma))
-            delta = max(delta,abs(V[s]-v))
-        if delta < theta:
-            break
-
-    print(V[0])
-

### 2019-06-01 homeworks/hw1/main.py

@@ -1,125 +1,0 @@
-import numpy as np
-from collections import defaultdict
-import sys
-import environment as e
-
-
-def generate_episode(env, policy):
-    episode = []
-    state = env.reset()
-    while True:
-        action = policy[state]
-        next_state, reward, done, info = env.step(action)
-        episode.append((state, action, reward))
-        state = next_state
-        if done:
-            break
-    return episode
-
-
-def mc_prediction_q(env, num_episodes, policy, gamma=1.0):
-    # initialize empty dictionaries of arrays
-    returns_sum = defaultdict(lambda: np.zeros(env.num_actions))
-    N = defaultdict(lambda: np.zeros(env.num_actions))
-    Q = defaultdict(lambda: np.zeros(env.num_actions))
-    # loop over episodes
-    for i_episode in range(1, num_episodes+1):
-        # monitor progress
-        if i_episode % 1000 == 0:
-            print("\rEpisode {}/{}.".format(i_episode, num_episodes), end="")
-            sys.stdout.flush()
-
-        # generate an episode
-        episode = generate_episode(env, policy)
-        # obtain the states, actions, and rewards
-        states, actions, rewards = zip(*episode)
-        # prepare for discounting
-        discounts = np.array([gamma**i for i in range(len(rewards)+1)])
-        # calculate and store the return for each visit in the episode
-        for i, (state, action, reward) in enumerate(episode):
-            returns_sum[state][action] += sum(rewards[i:]*discounts[:-(1+i)])
-            N[state][action] += 1
-
-    for state, values in returns_sum.items():
-        for action in range(env.num_actions):
-            Q[state][action] = values[action] / N[state][action]
-
-    return Q
-
-
-def generate_episode_from_Q(bj_env, Q, epsilon):
-    policy = np.ones(env.num_actions) / env.num_actions
-    episode = []
-    state = bj_env.reset()
-    while True:
-        if state in Q:
-            for action in range(env.num_actions):
-                if action == np.argmax(Q[state]):
-                    policy[action] = 1 - epsilon + epsilon / env.num_actions
-                else:
-                    policy[action] = epsilon / env.num_actions
-        else:
[…]

### 2019-06-01 homeworks/hw1/main2.py

@@ -1,54 +1,0 @@
-import numpy as np
-from collections import defaultdict
-import sys
-import environment as e
-
-
-def policy_risk(state):
-    return [0.5, 0.5]
-
-
-def policy_evaluation(env, policy, gamma=1, theta=1e-8):
-    V = defaultdict(lambda: 0)
-    i = 0
-    while True:
-        delta = 0
-        i += 1
-        for s in range(30): #len(env.mask)):
-            Vs = 0
-            for a, action_prob in enumerate(policy(s)):
-                for prob, next_state, reward, done in env.get_probs(s, a):
-                    Vs += action_prob * prob * (reward + gamma * V[next_state])
-
-            delta = max(delta, abs(Vs - V[s]))
-            V[s] = Vs
-
-        if delta < theta:
-            break
-
-    return V
-
-
-def q_from_v(env, V, s, gamma=1):
-    q = [0, 0]
-
-    for a in range(2):
-        for prob, next_state, reward, done in env.get_probs(s, a):
-            q[a] += prob * (reward + gamma * V[next_state])
-            if (s == 0) and (a == 0):
-                print(prob, next_state, reward, done)
-                print(q[a])
-
-    return q
-
-
-if __name__ == '__main__':
-    env = e.Environment([1, 1, 1, 0, 0, 0])
-    V = policy_evaluation(env, policy_risk, gamma=1, theta=1e-8)
-    print(V[0])
-    #Q = np.zeros([6, 2]) #defaultdict(lambda: [0, 0])
-    #for s in range(len(env.mask)):
-    #    Q[s] = q_from_v(env, V, s)
-
-    #print(Q)
-

### 2019-06-01 homeworks/hw2/final_solution.py

@@ -1,46 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-from scipy.optimize import fsolve
-
-
-if __name__ == '__main__':
-
-    #p = 0.24
-    #v = [0.0, 0, 13.7, 0.5, -4.7, 17.6, 0.0]
-    #r = [-1.2, -1.1, 4.8, 0, -1.7, 8.3, 2.2]
-
-    p = 0.62
-    v = [0.0, 0.0, 2.6, 19.5, 4.8, 0.6, 0.0]
-    r = [-2.8, 4.4, -1.1, 1.1, -2.2, 2.7, 1.6]
-
-    # Test
-    #p = 0.5
-    #v = [0, 3, 8, 2, 1, 2, 0]
-    #r = [0, 0, 0, 4, 1, 1, 1]
-
-    E = np.zeros(7)
-    E[1] = p * (r[0] + v[1]) + (1 - p) * (r[1] + v[2])
-    E[2] = p * (r[0] + r[2]) + (1 - p) * (r[1] + r[3]) + v[3]
-    E[3] = p * (r[0] + r[2]) + (1 - p) * (r[1] + r[3]) + r[4] + v[4]
-    E[4] = p * (r[0] + r[2]) + (1 - p) * (r[1] + r[3]) + r[4] + r[5] + v[5]
-    E[5] = p * (r[0] + r[2]) + (1 - p) * (r[1] + r[3]) + r[4] + r[5] + r[6] + v[6]
-    E[6] = p * (r[0] + r[2]) + (1 - p) * (r[1] + r[3]) + r[4] + r[5] + r[6] + 0 + 0
-    print(f'E: {E}')
-
-    def td(l):
-        return E[1] * (1 - l) + E[2] * l * (1 - l) + E[3] * l**2 * (1 - l) + E[4] * l**3 * (1 - l) + E[5] * l**4 * (1 - l) + E[6] * l**5 * (1 - l) + l**6 * E[6]
-
-    print(f'TD(0): {td(0)}')
-    print(f'TD(1): {td(1)}')
-
-    sol = fsolve(lambda l: td(l) - td(1), np.array([0.1]))[0]
-    print(f'Lambda: {sol:0.6f}')
-    print(f'Check: {td(sol)}')
-
-    plot = False
-    if plot:
-        x = np.linspace(0, 1)
-        plt.plot(x, td(x))
-        plt.axhline(td(1), color='red')
-        plt.show()
-

### 2019-06-01 homeworks/hw2/main2.py

@@ -1,51 +1,0 @@
-from scipy.optimize import fsolve,leastsq
-import numpy as np
-
-
-class TD_lambda:
-    def __init__(self, probToState,valueEstimates,rewards):
-        self.probToState = probToState
-        self.valueEstimates = valueEstimates
-        self.rewards = rewards
-        self.td1 = self.get_vs0(1)
-
-    def get_vs0(self, lambda_):
-        probToState = self.probToState
-        valueEstimates = self.valueEstimates
-        rewards = self.rewards
-        vs = dict(zip(['vs0','vs1','vs2','vs3','vs4','vs5','vs6'],list(valueEstimates)))
-
-        vs5 = vs['vs5'] + 1 * (rewards[6] + 1 * vs['vs6'] - vs['vs5'])
-
-        vs4 = vs['vs4'] + 1 * (rewards[5] + lambda_ * rewards[6] + lambda_ * vs['vs6'] + (1 - lambda_) * vs['vs5'] - vs['vs4'])
-
-        vs3 = vs['vs3'] + 1 * (rewards[4] + lambda_ * rewards[5] + lambda_**2 * rewards[6] + lambda_**2 * vs['vs6'] + lambda_ * (1 - lambda_) * vs['vs5'] + (1 - lambda_) * vs['vs4'] - vs['vs3'])
-
-        vs1 = vs['vs1'] + 1*(rewards[2]+lambda_*rewards[4]+lambda_**2*rewards[5]+lambda_**3*rewards[6]+lambda_**3*vs['vs6']+lambda_**2*(1-lambda_)*vs['vs5']+lambda_*(1-lambda_)*vs['vs4']+ \
-                             (1-lambda_)*vs['vs3']-vs['vs1'])
-
-        vs2 = vs['vs2'] + 1*(rewards[3]+lambda_*rewards[4]+lambda_**2*rewards[5]+lambda_**3*rewards[6]+lambda_**3*vs['vs6']+lambda_**2*(1-lambda_)*vs['vs5']+lambda_*(1-lambda_)*vs['vs4']+ \
-                             (1-lambda_)*vs['vs3']-vs['vs2'])
-
[…]

### 2019-06-01 project_1/backup/final_solution.py
Removing unnecessary files

@@ -1,565 +1,0 @@
-# -*- coding: utf-8 -*-
-"""OMSCS Reinforcement Learning - CS-7642-O03 - Project #1 Solution
-
-This code solves Project #1 from OMSCS Reinforcement Learning - CS-7642-O03,
-generating the charts used in the report. Running it is straightforward:
-
-    $ python final_solution.py
-
-This will create the figure with all 3 charts in the file 'images/figure.png',
-used in the report. It also creates them in individual files,
-'images/figure1.png', 'images/figure2.png', and 'images/figure3.png', which are
-replications of Sutton's Figure 3, 4 and 5 respectively.
-
-Todo:
-    * Write report
-    * Write README.md
-    * Test in ubuntu
-    * Include a requirements.txt in the package
-    * Create folder 'images' if it doesn't exist
-
-Created by Carlos Souza (souza@gatech.edu). May-2019.
-
-"""
-
-import numpy as np
-import pandas as pd
-#import matplotlib
-#matplotlib.use('Agg')
-import matplotlib.pyplot as plt
-from tqdm import tqdm
-import random
-from collections import defaultdict
-from time import sleep
-
-
-class Environment(object):
-    """Environment class.
-
-    This class simulates the Random Walk environment. At initialization, it
-    receives the random walk size. Then, at each time step, it updates the
-    current state and returns the experience tuples with next state, reward, and
-    whether it reached a terminal state. Inspired in OpenAI gym environments.
-
-    Args:
-        size (int): Size of the random walk, excluding terminals states. For
-            example, in Sutton's original paper, random walk has size = 5.
-
-    Attributes:
-        size (int): Size of the random walk, including terminal states.
-        current_state (int): Index representing the current state. For example:
-            if the random walk has size 7 (including terminal states), in the
-            beginning of the simulation, when in center, current_state is 3
-            (middle point between 0 and 6). If at next time step the move is to
-            the right, current_index become 4. If left, current_index become 2.
-            In this example, current_state = 0 is the terminal state at extreme
-            left (zero reward) and current_state = 6 is the terminal state at
-            extreme right (reward +1).
-        t (int): Number of time steps elapsed since beginning of simulation.
-
-    """
-    def __init__(self, size=5):
-        self.size = size + 2  # Adding 2 terminal states at beginning and end
-        self.current_state = None
-        self.t = None
-        self.reset()
-
-    def reset(self):
-        """Reset method.
-
[…]

### 2019-06-01 project_1/backup/main.py

@@ -1,240 +1,0 @@
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-from collections import defaultdict, deque
-from tqdm import tqdm
-import random
-
-
-class Environment(object):
-    def __init__(self, size=5):
-        self.size = size + 2  # Adding 2 terminal states at beginning and end
-        self.current_state = None
-        self.t = None
-        self.reset()
-
-    def reset(self):
-        self.current_state = int((self.size - 1) / 2)
-        self.t = 0
-        return self.current_state
-
-    def step(self):
-        self.t += 1
-        next_state = np.random.choice([-1, 1]) + self.current_state
-        self.current_state = next_state
-        if next_state == self.size - 1:
-            reward = 1
-        else:
-            reward = 0
-
-        if next_state == self.size - 1 or next_state == 0:
-            done = True
-        else:
-            done = False
-
-        return next_state, reward, done, self.t
-
-
-class Solver(object):
-    def __init__(self, env=None, size=5):
-        self.env = Environment(size=size) if env is None else env
-        self.v_true = np.zeros(self.env.size)
-        for i in range(self.env.size):
-            self.v_true[i] = i / (self.env.size - 1)
-
-    def estimate_v_td0(self, v0=None, num_episodes=100, alpha=0.05, gamma=1.0):
-        v = np.repeat(0.5, self.env.size) if v0 is None else v0
-        v[0] = 0
-        v[self.env.size - 1] = 0
-        rms = np.zeros(num_episodes)
-
-        for i_episode in range(num_episodes):
-            state = self.env.reset()
-            done = False
-            while not done:
-                next_state, reward, done, info = self.env.step()
-                v[state] += alpha * (reward + gamma * v[next_state] - v[state])
-                state = next_state
-
-            error = self.v_true[1:-1] - v[1:-1]
-            rms[i_episode] = np.sqrt(np.mean(error**2))
-
-        return v[1:-1], rms
-
-    def estimate_v_mc(self, v0=None, num_episodes=100, gamma=1.0):
-        v = np.repeat(0.5, self.env.size) if v0 is None else v0
-        v[0] = 0
-        v[self.env.size - 1] = 0
-        rms = np.zeros(num_episodes)
-        returns = defaultdict(list)
-
-        for i_episode in range(num_episodes):
-            state = self.env.reset()
-            done = False
-            experience = []
[…]

### 2019-06-01 project_1/backup/main2.py

@@ -1,112 +1,0 @@
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-from collections import defaultdict, deque
-from tqdm import tqdm
-import random
-
-
-class Environment(object):
-    def __init__(self, size=5):
-        self.s = size
-        self.v_true = np.zeros(self.s)
-        for i in range(self.s):
-            self.v_true[i] = (i + 1) / (self.s + 1)
-
-    def generate_data(self, num_sequences=10, num_training_sets=100):
-        training_sets = []
-        for i_training_set in range(num_training_sets):
-            sequences = []
-            for i_sequence in range(num_sequences):
-                state = np.zeros(self.s)
-                curr_index = int((self.s - 1) / 2)
-                state[curr_index] = 1
-                done = False
-                sequence = [state.copy()]
-                while not done:
-                    next_index = np.random.choice([curr_index - 1, curr_index + 1])
-                    if next_index < 0:
-                        sequence.append(0)
-                        break
-                    elif next_index >= self.s:
-                        sequence.append(1)
-                        break
-                    else:
-                        state[curr_index] = 0
-                        state[next_index] = 1
-                        sequence.append(state.copy())
-                        curr_index = next_index
-
-                sequences.append(sequence)
-
-            training_sets.append(sequences)
-
-        return training_sets
-
-    def print_data(self, training_sets):
-        for s, sequences in enumerate(training_sets):
-            print(f'TRAINING SET {s + 1}' + ('*' * 100))
-            for i, sequence in enumerate(sequences):
-                print(f'Sequence {i + 1}')
-                for j, step in enumerate(sequence):
-                    print(f'Step {j + 1}: {step}')
-
-                print(' ')
-
-    def estimate_v_tdlambda(self, lambda_, training_set, alpha=0.1, gamma=1.0):
-        v = np.repeat(0.5, self.s)
-        rms = np.zeros(len(training_set))
-
-        for i, sequence in enumerate(training_set):
-            state = sequence[0]
-            curr_index = np.where(state == 1)
-            #eligibility = np.zeros(self.s)
-            for j, next_state in enumerate(sequence[1:]):
-                if type(next_state) is np.ndarray:
-                    reward = 0
-                    next_index = np.where(next_state == 1)
[…]

### 2019-06-01 project_1/paper.tex
Reorganizing folders to commit to GT server
Adding to GT server

--- 2019-05-29 project1/paper.tex
@@
 \author{
-  Carlos Souza\\
+  Carlos Souza\thanks{Latest Git hash: DO NOT FORGET TO INCLUDE LATEST GIT HASH}\\
   %Department of Computer Science\\
@@
 \keywords{Reinforcement learning \and Temporal-difference (TD) learning \and Markov reward process}
+%\githash{DO NOT FORGET TO INCLUDE LATEST GIT HASH}
 
@@
 A bounded random walk is a Markov Reward Process.
-The episodes always start in the middle state $D$.
+The episodes always start in the middle state $C$.
 At each time step, the walk moves to the neighbor state, either left or right, with equal probability.
-The episode ends when walk reaches either extreme states $A$ or $G$.
-If termination happens by reaching the extreme right ($G$), environment provides a reward of +1;
+The episode ends when walk reaches either extreme states ($A$'s left or $E$'s right).
+If termination happens by reaching the extreme right, environment provides a reward of +1;
 all the other rewards are zero.
@@
 \begin{figure}[t]
-    \includegraphics[scale=0.3]{./images/random_walk.eps}
+    \includegraphics[scale=0.3]{./images/random_walk2.eps}
     \centering
     \caption{A bounded random walk generator used in Sutton's original paper.
-    All episodes start in state $D$.
+    All episodes start in state $C$.
     At each time step, walk has equal probability of either moving left or right.
-    Episode ends when walk reaches either $A$ or $G$.
-    In case of episode terminating in $G$, +1 reward is received;
+    Episode ends when walk reaches either extreme left or right.
+    In case of episode terminating in extreme right, +1 reward is received;
     otherwise, reward is zero.}
@@
 
-A typical episode terminating in the extreme right might be the following state-reward sequence: $D, 0, C, 0, D, 0, E, 0, F, 0, G, 1$.
-An example of a walk ending in the extreme left might be: $D, 0, C, 0, B, 0, A, 0$.
+A typical episode terminating in the extreme right might be the following state-reward sequence: C, 0, B, 0, C, 0, D, 0, E, 0, extreme-right (terminal), 1.
+An example of a walk ending in the extreme left might be: C, 0, B, 0, A, 0, extreme-left (terminal), 0.
 This MRP is undiscounted ($\gamma = 1$);
 therefore, the true value of each state is the probability of terminating on the extreme right by starting from that specific state.
-Thus, the true values from all states, from $A$ to $G$ respectively, are
+Thus, the true values from all states, from $A$ to $E$ respectively, are
 
 \begin{equation}
-    v_{true} = \Bigg[0, \frac{1}{6}, \frac{2}{6}, \frac{3}{6}, \frac{4}{6}, \frac{5}{6}, 1\Bigg]
+    v_{true} = \Bigg[\frac{1}{6}, \frac{2}{6}, \frac{3}{6}, \frac{4}{6}, \frac{5}{6}\Bigg]
     \centering

### 2019-06-01 project_1/final_solution.py

@@ -11,10 +11,6 @@ used in the report. It also creates them in individual files,
 'images/figure1.png', 'images/figure2.png', and 'images/figure3.png', which are
 replications of Sutton's Figure 3, 4 and 5 respectively.
 
-TODO:
-    * Adjust random_walk illustration and text(from A to E, to be compatible
-      with video screenshot)
-
 Created by Carlos Souza (souza@gatech.edu)
 May-2019
 

### 2019-06-03 homeworks/hw1/environment.py
Adding branch homeworks, different from master

@@ -1,0 +1,99 @@
+import numpy as np
+from random import randint
+from numpy.random import choice
+import pprint
+
+
+class Environment(object):
+	def __init__(self, mask):
+		self.mask = mask
+		self.bankroll = 0
+		self.count = 0
+		self.num_actions = 2
+
+	def __str__(self):
+		return f'N = {len(self.mask)}, isBadSide = {self.mask}'
+		
+	def step(self, action):
+		"""Step function
+
+		Args:
+			action: The action being taken
+
+		Returns:
+			tuple: Containing next state, reward, done and info
+
+		"""
+		self.count += 1
+		if action == 0:
+			return self.bankroll, 0, True, (self.count, None)
+		
+		roll = randint(1, len(self.mask))
+		if self.mask[roll - 1] == 1:
+			return 0, -self.bankroll, True, (self.count, roll)
+		else:
+			self.bankroll += roll
+			return self.bankroll, roll, False, (self.count, roll)
+		
+	def reset(self):
+		self.bankroll = 0
+		self.count = 0
+		return self.bankroll
+
+	def get_probs(self, state, action):
+		"""Get transition probabilities after taking action in a given state
+
+		Args:
+			state: The current state
+			action: The action taken
+
+		Returns:
+			list of tuples: A list of tuples containing probability, next_state,
+				reward and done
+
+		"""
+		#if self.mask[state - 1] == 1:
+		#	return [(1, 0, -state, True)]
+
+		if action == 0:
+			return [(1, state, 0, True)]
+
+		result = []
+		for roll in range(1, len(self.mask) + 1):
+			if self.mask[roll - 1] == 1:
+				leaf = (1.0 / len(self.mask), 1000, -state, True)
+			else:
+				leaf = (1.0 / len(self.mask), state + roll, roll, False)
+
+			result.append(leaf)
+
+		return result
+
+
+def q_from_v(env, V, s, gamma=1):
+	q = np.zeros(2)
+	for a in range(2):
+		for prob, next_state, reward, done in env.P[s][a]:
+			q[a] += prob * (reward + gamma * V[next_state])
+	return q
+
+		
+if __name__ == '__main__':
+	pp = pprint.PrettyPrinter()
+	env = Environment([1, 1, 1, 0, 0, 0])
+	pp.pprint(env.get_probs(4, 1))
+
+	exit(0)
+
+	for i_episode in range(20):
+		print(f'Episode {i_episode + 1}')
+		state = env.reset()
[…]

### 2019-06-03 homeworks/hw1/final_solution.py

@@ -1,0 +1,76 @@
+import numpy as np
+from collections import defaultdict
+
+
+def get_probs(mask, state, action):
+    """Get transition probabilities after taking action in a given state
+
+    Args:
+        state: The current state
+        action: The action taken
+
+    Returns:
+        list of tuples: A list of tuples containing probability, next_state,
+            reward and done
+
+    """
+    if action == 0:
+        return [(1, state, 0, True)]
+
+    result = []
+    for roll in range(1, len(mask) + 1):
+        if mask[roll - 1] == 1:
+            leaf = (1.0 / len(mask), 1000, -state, True)
+        else:
+            leaf = (1.0 / len(mask), state + roll, roll, False)
+
+        result.append(leaf)
+
+    return result
+
+
+def get_possible_states(mask, origin=[0]):
+    result = []
+    for node in origin:
+        for i in range(1, len(mask) + 1):
+            if (mask[i - 1] == 0) and ((node + i) not in result) and ((node + i) not in origin):
+                result.append(node + i)
+
+    return origin + result
+
+
+def q_from_v(mask, V, s, gamma=1):
+    q = np.zeros(2)
+    for a in range(2):
+        for prob, next_state, reward, done in get_probs(mask, s, a):
+            q[a] += prob * (reward + gamma * V[next_state])
+    return q
+
+
+if __name__ == '__main__':
+    #mask = [1, 1, 1, 0, 0, 0]
+    #mask = [1,1,1,1,1,1,0,1,0,1,1,0,1,0,1,0,0,1,0,0,1,0]
+    #mask = [1,1,1,1,0,0,0,0,1,0,1,0,1,1,0,1,0,0,0,1,0]
+    mask = [0,0,0,0,0,0,0,0,0,1,0,0,1,1,0,0]
+    previous = [0]
+    for i in range(4):
+        states = get_possible_states(mask, origin=previous)
+        previous = states
+
+    states += [1000]
+    print(states)
+
+    gamma = 1
+    theta = 1e-8
+    V = defaultdict(lambda: 0) #np.zeros(len(states))
+    while True:
+        delta = 0
+        for s in states:
+            v = V[s]
+            V[s] = max(q_from_v(mask, V, s, gamma))
+            delta = max(delta,abs(V[s]-v))
+        if delta < theta:
+            break
+
+    print(V[0])
+

### 2019-06-03 homeworks/hw1/main2.py

@@ -1,0 +1,54 @@
+import numpy as np
+from collections import defaultdict
+import sys
+import environment as e
+
+
+def policy_risk(state):
+    return [0.5, 0.5]
+
+
+def policy_evaluation(env, policy, gamma=1, theta=1e-8):
+    V = defaultdict(lambda: 0)
+    i = 0
+    while True:
+        delta = 0
+        i += 1
+        for s in range(30): #len(env.mask)):
+            Vs = 0
+            for a, action_prob in enumerate(policy(s)):
+                for prob, next_state, reward, done in env.get_probs(s, a):
+                    Vs += action_prob * prob * (reward + gamma * V[next_state])
+
+            delta = max(delta, abs(Vs - V[s]))
+            V[s] = Vs
+
+        if delta < theta:
+            break
+
+    return V
+
+
+def q_from_v(env, V, s, gamma=1):
+    q = [0, 0]
+
+    for a in range(2):
+        for prob, next_state, reward, done in env.get_probs(s, a):
+            q[a] += prob * (reward + gamma * V[next_state])
+            if (s == 0) and (a == 0):
+                print(prob, next_state, reward, done)
+                print(q[a])
+
+    return q
+
+
+if __name__ == '__main__':
+    env = e.Environment([1, 1, 1, 0, 0, 0])
+    V = policy_evaluation(env, policy_risk, gamma=1, theta=1e-8)
+    print(V[0])
+    #Q = np.zeros([6, 2]) #defaultdict(lambda: [0, 0])
+    #for s in range(len(env.mask)):
+    #    Q[s] = q_from_v(env, V, s)
+
+    #print(Q)
+

### 2019-06-03 homeworks/hw2/final_solution.py

@@ -1,0 +1,46 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.optimize import fsolve
+
+
+if __name__ == '__main__':
+
+    #p = 0.24
+    #v = [0.0, 0, 13.7, 0.5, -4.7, 17.6, 0.0]
+    #r = [-1.2, -1.1, 4.8, 0, -1.7, 8.3, 2.2]
+
+    p = 0.62
+    v = [0.0, 0.0, 2.6, 19.5, 4.8, 0.6, 0.0]
+    r = [-2.8, 4.4, -1.1, 1.1, -2.2, 2.7, 1.6]
+
+    # Test
+    #p = 0.5
+    #v = [0, 3, 8, 2, 1, 2, 0]
+    #r = [0, 0, 0, 4, 1, 1, 1]
+
+    E = np.zeros(7)
+    E[1] = p * (r[0] + v[1]) + (1 - p) * (r[1] + v[2])
+    E[2] = p * (r[0] + r[2]) + (1 - p) * (r[1] + r[3]) + v[3]
+    E[3] = p * (r[0] + r[2]) + (1 - p) * (r[1] + r[3]) + r[4] + v[4]
+    E[4] = p * (r[0] + r[2]) + (1 - p) * (r[1] + r[3]) + r[4] + r[5] + v[5]
+    E[5] = p * (r[0] + r[2]) + (1 - p) * (r[1] + r[3]) + r[4] + r[5] + r[6] + v[6]
+    E[6] = p * (r[0] + r[2]) + (1 - p) * (r[1] + r[3]) + r[4] + r[5] + r[6] + 0 + 0
+    print(f'E: {E}')
+
+    def td(l):
+        return E[1] * (1 - l) + E[2] * l * (1 - l) + E[3] * l**2 * (1 - l) + E[4] * l**3 * (1 - l) + E[5] * l**4 * (1 - l) + E[6] * l**5 * (1 - l) + l**6 * E[6]
+
+    print(f'TD(0): {td(0)}')
+    print(f'TD(1): {td(1)}')
+
+    sol = fsolve(lambda l: td(l) - td(1), np.array([0.1]))[0]
+    print(f'Lambda: {sol:0.6f}')
+    print(f'Check: {td(sol)}')
+
+    plot = False
+    if plot:
+        x = np.linspace(0, 1)
+        plt.plot(x, td(x))
+        plt.axhline(td(1), color='red')
+        plt.show()
+

### 2019-06-03 homeworks/hw2/main2.py

@@ -1,0 +1,51 @@
+from scipy.optimize import fsolve,leastsq
+import numpy as np
+
+
+class TD_lambda:
+    def __init__(self, probToState,valueEstimates,rewards):
+        self.probToState = probToState
+        self.valueEstimates = valueEstimates
+        self.rewards = rewards
+        self.td1 = self.get_vs0(1)
+
+    def get_vs0(self, lambda_):
+        probToState = self.probToState
+        valueEstimates = self.valueEstimates
+        rewards = self.rewards
+        vs = dict(zip(['vs0','vs1','vs2','vs3','vs4','vs5','vs6'],list(valueEstimates)))
+
+        vs5 = vs['vs5'] + 1 * (rewards[6] + 1 * vs['vs6'] - vs['vs5'])
+
+        vs4 = vs['vs4'] + 1 * (rewards[5] + lambda_ * rewards[6] + lambda_ * vs['vs6'] + (1 - lambda_) * vs['vs5'] - vs['vs4'])
+
+        vs3 = vs['vs3'] + 1 * (rewards[4] + lambda_ * rewards[5] + lambda_**2 * rewards[6] + lambda_**2 * vs['vs6'] + lambda_ * (1 - lambda_) * vs['vs5'] + (1 - lambda_) * vs['vs4'] - vs['vs3'])
+
+        vs1 = vs['vs1'] + 1*(rewards[2]+lambda_*rewards[4]+lambda_**2*rewards[5]+lambda_**3*rewards[6]+lambda_**3*vs['vs6']+lambda_**2*(1-lambda_)*vs['vs5']+lambda_*(1-lambda_)*vs['vs4']+ \
+                             (1-lambda_)*vs['vs3']-vs['vs1'])
+
+        vs2 = vs['vs2'] + 1*(rewards[3]+lambda_*rewards[4]+lambda_**2*rewards[5]+lambda_**3*rewards[6]+lambda_**3*vs['vs6']+lambda_**2*(1-lambda_)*vs['vs5']+lambda_*(1-lambda_)*vs['vs4']+ \
+                             (1-lambda_)*vs['vs3']-vs['vs2'])
+
[…]

### 2019-06-03 homeworks/hw3/hw3-tester/hw3_tester.py

@@ -1,0 +1,313 @@
+import os
+import json
+import math
+import argparse
+import mdptoolbox
+import numpy as np
+import progressbar
+import logging as log
+from decimal import *
+
+
+class MDP:
+    """Convert the plain object description of the mdp into gamma and T and R matrices"""
+    def __init__(self, descr):
+        self.descr = descr
+        self.gamma = descr["gamma"]
+        self.nS = len(descr["states"])
+        self.nA = len(descr["states"][0]["actions"])
+        self.transitions = np.zeros((self.nA, self.nS, self.nS))
+        self.rewards = np.zeros((self.nA, self.nS, self.nS))
+        state_indexes = {state["id"]: i for i, state in enumerate(descr["states"])}
+        for state in descr["states"]:
+            assert len(state["actions"]) == self.nA, "All states must have same number of possible actions"
+            for i, action in enumerate(state["actions"]):
+                for transition in action["transitions"]:
+                    state_index = state_indexes[state["id"]]
+                    new_state_index = state_indexes[transition["to"]]
+                    self.transitions[i, state_index, new_state_index] = transition["probability"]
+                    self.rewards[i, state_index, new_state_index] = transition["reward"]
+
+def get_iterations_with_mdptoolbox(mdp_descr):
+    log.info('in get_iterations function')
+    mdp = MDP(mdp_descr)
+
+    log.info('running policy improvement')
+    # (policy, v, it)
+    results = []
+    bar = progressbar.ProgressBar(maxval=1000,
+                                  widgets=[progressbar.Bar('=', '[', ']'), ' ',
+                                           progressbar.Percentage()])
+    bar.start()
+    for t in range(1000):
+        np.seterr(all='raise')
+        try:
+            initial_policy = np.random.choice(mdp.nA, size=mdp.nS)
+            pi = mdptoolbox.mdp.PolicyIteration(
+                mdp.transitions,
+                mdp.rewards,
+                mdp.gamma,
+                policy0=initial_policy,
+                eval_type=1
+            )
+            pi.setSilent()
+            pi.run()
+            result = pi.iter
+            results.append(result)
+        except Exception as e:
+            log.error('exception: ' + e.message)
+            log.error('won\'t count trial')
+        bar.update(t + 1)
+    bar.finish()
+    if len(results) == 0:
+        log.critical('empty results, please check for errors')
+        exit(-2)
+    results = np.array(results)
[…]

### 2019-06-04 homeworks/hw3/main.py
HW3 first commit

@@ -1,0 +1,15 @@
+import numpy as np
+import json
+
+if __name__ == '__main__':
+    d = dict()
+    d['gamma'] = 0.75
+    d['states'] = list()
+    d['states'].append('a')
+    d['states'].append('b')
+    d['states'].append('c')
+    print(d)
+
+    j = json.dumps(d, indent=4)
+    print(j)
+

### 2019-06-24 project_2/agent_pytorch.py
Keep agents learning after convergence

@@ -279,6 +279,7 @@ class Agent(object):
                                           'average', 'avg_q_values'])
         t_start = time()
         best_score = float('-inf')
+        self.epsilon = 0
 
         for i_episode in range(max_episodes):
             t = time()

### 2019-06-24 project_2/main_pytorch.py

@@ -28,7 +28,7 @@ if __name__ == '__main__':
         return max(min_epsilon, curr_epsilon * decay)
 
     def epsilon_decay2(curr_epsilon, i_episode, min_epsilon=0.05):
-        return max(min_epsilon, np.exp(-i_episode / 70))
+        return max(min_epsilon, np.exp(-i_episode / 40))
 
     print('Agent parameters: ')
     print(vars(agent))

### 2019-06-24 project_2/playground.py

@@ -22,11 +22,11 @@ y8 = 1 / (1 + np.exp(x/40 - 6))
 y9 = 1 / (1 + np.exp(x/3 - 6))
 
 fig, ax = plt.subplots()
-ax.plot(x, y0)
-ax.plot(x, y1)
-ax.plot(x, y2)
-ax.plot(x, y3)
-ax.plot(x, y4)
+ax.plot(x, y0, label='y0')
+ax.plot(x, y1, label='y1')
+ax.plot(x, y2, label='y2')
+ax.plot(x, y3, label='y3')
+ax.plot(x, y4, label='y4')
 
 #ax.plot(x, y5)
 #ax.plot(x, y6)
@@ -35,4 +35,6 @@ ax.plot(x, y4)
 #ax.plot(x, y9)
 
 ax.grid()
+ax.legend()
+
 plt.show()

### 2019-06-24 project_2/parallel_exp2.py
Fixing bug on parallel experiments

@@ -1,0 +1,97 @@
+import pandas as pd
+import numpy as np
+import os
+import json
+import multiprocessing as mp
+from time import time
+from datetime import timedelta
+import argparse
+import agent_pytorch as ag
+from epsilon_curves import get_epsilon_decay
+
+
+def run_experiment(protocol, experiment, folder):
+    name = mp.current_process().name
+
+    decay_function = experiment['decay_function'] if 'decay_function' in \
+                                                     experiment else 0
+    epsilon_decay = get_epsilon_decay(decay_function=decay_function)
+
+    t_run_start = time()
+    print(f'Starting {name}/{protocol["global_params"]["runs_per_experiment"]}, '
+          f'experiment {experiment["id"]} ({i + 1}/{len(protocol["experiments"])})')
+
+    agent = ag.Agent(batch_size=protocol['global_params']['batch_size'],
+                     layers=protocol['global_params']['layers'],
+                     alpha=experiment['alpha'],
+                     gamma=experiment['gamma'],
+                     dueling=experiment['dueling'],
+                     double=experiment['double'],
+                     prioritized_er=experiment['prioritized_er'])
+
+    df_scores = agent.train(run_n=name,
+                            max_episodes=protocol['global_params']['max_episodes'],
+                            stop_when_solved=protocol['global_params']['stop_when_solved'],
+                            keep_learning=protocol['global_params']['keep_learning'],
+                            epsilon_decay=epsilon_decay,
+                            print_frequency=protocol['global_params']['print_frequency'],
+                            render=False,
+                            print_same_line=False,
+                            log_floydhub=False)
+    run_filename = f'{folder}/df_{eid}_run_{j + 1:02d}.csv'
+    df_scores.to_csv(run_filename, index=False)
+
+    if 'test_for' in protocol['global_params']:
+        df_scores_test = agent.test(run_n=name,
+                                    max_episodes=protocol['global_params']['test_for'],
+                                    print_frequency=protocol['global_params']['print_frequency'],
+                                    render=False,
+                                    print_same_line=False)
+        test_run_filename = f'{folder}/df_{eid}_run_{j + 1:02d}_test.csv'
+        df_scores_test.to_csv(test_run_filename, index=False)
+
+    print(f'{name}\tTime to complete: {timedelta(seconds=time() - t_run_start)}\n')
+
+
+if __name__ == '__main__':
[…]

### 2019-06-24 project_2/parallel_experiments.py

@@ -10,7 +10,7 @@ import agent_pytorch as ag
 from epsilon_curves import get_epsilon_decay
 
 
-def run_experiment(protocol, experiment, folder):
+def run_experiment(protocol, experiment, run_filename, test_run_filename):
     name = mp.current_process().name
 
     decay_function = experiment['decay_function'] if 'decay_function' in \
@@ -38,7 +38,7 @@ def run_experiment(protocol, experiment, folder):
                             render=False,
                             print_same_line=False,
                             log_floydhub=False)
-    run_filename = f'{folder}/df_{eid}_run_{j + 1:02d}.csv'
+
     df_scores.to_csv(run_filename, index=False)
 
     if 'test_for' in protocol['global_params']:
@@ -47,7 +47,6 @@ def run_experiment(protocol, experiment, folder):
                                     print_frequency=protocol['global_params']['print_frequency'],
                                     render=False,
                                     print_same_line=False)
-        test_run_filename = f'{folder}/df_{eid}_run_{j + 1:02d}_test.csv'
         df_scores_test.to_csv(test_run_filename, index=False)
 
     print(f'{name}\tTime to complete: {timedelta(seconds=time() - t_run_start)}\n')
@@ -106,8 +105,12 @@ if __name__ == '__main__':
         jobs = []
         eid = experiment['id']
         for j in range(protocol['global_params']['runs_per_experiment']):
+            run_filename = f'{full_experiment_folder}/df_{eid}_run_{j + 1:02d}.csv'
+            test_run_filename = f'{full_experiment_folder}/df_{eid}_run_{j + 1:02d}_test.csv'
+
             job = mp.Process(target=run_experiment, args=(protocol, experiment,
-                                                          full_experiment_folder),
+                                                          run_filename,
+                                                          test_run_filename),
                              name=f'run_{j + 1:02d}')
             jobs.append(job)
 

### 2019-06-25 project_2/agent_pytorch.py
Preparing code to do single-run experiments in parallel and parametrizing optimizer

@@ -16,8 +16,8 @@ from torch_networks import DQN, DuelingDQN
 class Agent(object):
     def __init__(self, gamma=0.99, alpha=0.0005, memory_capacity=10000,
                  batch_size=64, layers=[512, 512], dueling=True, double=True,
-                 prioritized_er=False, tau=1.0, min_epsilon=0.05, huber=False,
-                 disable_cuda=False):
+                 prioritized_er=False, optimizer='rmsprop', tau=1.0,
+                 min_epsilon=0.05, huber=False, disable_cuda=False):
         if not disable_cuda and torch.cuda.is_available():
             self.device = torch.device('cuda')
         else:
@@ -41,7 +41,17 @@ class Agent(object):
         self.min_epsilon = min_epsilon
 
         self.huber = huber
-        self.optimizer = torch.optim.RMSprop(self.Q.parameters(), lr=self.alpha)
+        if optimizer == 'rmsprop':
+            self.optimizer = torch.optim.RMSprop(self.Q.parameters(),
+                                                 lr=self.alpha)
+        elif optimizer == 'adam':
+            self.optimizer = torch.optim.Adam(self.Q.parameters(),
+                                              lr=self.alpha)
+        elif optimizer == 'sgd':
+            self.optimizer = torch.optim.SGD(self.Q.parameters(), lr=self.alpha)
+        else:
+            raise ValueError('Optimizer must be "rmsprop", "adam" or "sgd"')
+
         self.update_target_weights()
 
     def build_model(self, layers, dueling=True):

### 2019-06-25 project_2/parallel_experiments.py

@@ -10,7 +10,8 @@ import agent_pytorch as ag
 from epsilon_curves import get_epsilon_decay
 
 
-def run_experiment(protocol, experiment, run_filename, test_run_filename):
+def run_experiment(protocol, experiment, exp_filename, test_exp_filename,
+                   counter):
     name = mp.current_process().name
 
     decay_function = experiment['decay_function'] if 'decay_function' in \
@@ -18,10 +19,10 @@ def run_experiment(protocol, experiment, run_filename, test_run_filename):
     epsilon_decay = get_epsilon_decay(decay_function=decay_function)
 
     t_run_start = time()
-    print(f'Starting {name}/{protocol["global_params"]["runs_per_experiment"]}, '
-          f'experiment {experiment["id"]} ({i + 1}/{len(protocol["experiments"])})')
+    print(f'Starting experiment {experiment["id"]} ({counter + 1}/{len(protocol["experiments"])})')
 
-    agent = ag.Agent(batch_size=protocol['global_params']['batch_size'],
+    agent = ag.Agent(optimizer=protocol['global_params']['optimizer'],
+                     batch_size=protocol['global_params']['batch_size'],
                      layers=protocol['global_params']['layers'],
                      alpha=experiment['alpha'],
                      gamma=experiment['gamma'],
@@ -38,8 +39,7 @@ def run_experiment(protocol, experiment, run_filename, test_run_filename):
                             render=False,
                             print_same_line=False,
                             log_floydhub=False)
-
-    df_scores.to_csv(run_filename, index=False)
+    df_scores.to_csv(exp_filename, index=False)
 
     if 'test_for' in protocol['global_params']:
         df_scores_test = agent.test(run_n=name,
@@ -47,32 +47,9 @@ def run_experiment(protocol, experiment, run_filename, test_run_filename):
                                     print_frequency=protocol['global_params']['print_frequency'],
                                     render=False,
                                     print_same_line=False)
-        df_scores_test.to_csv(test_run_filename, index=False)
+        df_scores_test.to_csv(test_exp_filename, index=False)
 
     print(f'{name}\tTime to complete: {timedelta(seconds=time() - t_run_start)}\n')
-
-
-def combine_experiment_runs(experiment_folder, experiment_id, runs, df_filename):
-    def load_experiment_df(filename, suffix):
-        return pd.read_csv(filename, dtype={'episode': int},
[…]

### 2019-06-25 project_2/parallel_runs.py

@@ -10,7 +10,8 @@ import agent_pytorch as ag
 from epsilon_curves import get_epsilon_decay
 
 
-def run_experiment(protocol, experiment, folder):
+def run_experiment(protocol, experiment, run_filename, test_run_filename,
+                   counter):
     name = mp.current_process().name
 
     decay_function = experiment['decay_function'] if 'decay_function' in \
@@ -19,7 +20,7 @@ def run_experiment(protocol, experiment, folder):
 
     t_run_start = time()
     print(f'Starting {name}/{protocol["global_params"]["runs_per_experiment"]}, '
-          f'experiment {experiment["id"]} ({i + 1}/{len(protocol["experiments"])})')
+          f'experiment {experiment["id"]} ({counter + 1}/{len(protocol["experiments"])})')
 
     agent = ag.Agent(batch_size=protocol['global_params']['batch_size'],
                      layers=protocol['global_params']['layers'],
@@ -38,7 +39,7 @@ def run_experiment(protocol, experiment, folder):
                             render=False,
                             print_same_line=False,
                             log_floydhub=False)
-    run_filename = f'{folder}/df_{eid}_run_{j + 1:02d}.csv'
+
     df_scores.to_csv(run_filename, index=False)
 
     if 'test_for' in protocol['global_params']:
@@ -47,10 +48,32 @@ def run_experiment(protocol, experiment, folder):
                                     print_frequency=protocol['global_params']['print_frequency'],
                                     render=False,
                                     print_same_line=False)
-        test_run_filename = f'{folder}/df_{eid}_run_{j + 1:02d}_test.csv'
         df_scores_test.to_csv(test_run_filename, index=False)
 
     print(f'{name}\tTime to complete: {timedelta(seconds=time() - t_run_start)}\n')
+
+
+def combine_experiment_runs(experiment_folder, experiment_id, runs, df_filename):
+    def load_experiment_df(filename, suffix):
+        return pd.read_csv(filename, dtype={'episode': int},
+                           index_col=0).add_suffix(suffix)
+
+    frames = [load_experiment_df(f'{experiment_folder}/df_{experiment_id}_run_{x + 1:02d}.csv',
+                                 suffix=f'_{x + 1:02d}') for x in range(runs)]
+    df = pd.concat(frames, axis=1)
+
+    def column_list(variable, runs):
+        return [f'{variable}_{x + 1:02d}' for x in range(runs)]
+
+    columns = []
+    for variable in ['score', 'average', 'avg_q_values']:
[…]

### 2019-06-25 project_2/playground.py

@@ -2,39 +2,53 @@ import numpy as np
 import matplotlib.pyplot as plt
 import os
 
-#rank = np.arange(0, 17)
-#k = rank * 10 + 1
-#print(rank)
-#print(k)
+
+def exp_curves():
+    x = np.linspace(0, 500, 100)
+    y0 = np.exp(-x / 70)
+    y1 = np.exp(-x / 5)
+    y2 = np.exp(-x / 20)
+    y3 = np.exp(-x / 40)
+    y4 = np.exp(-x / 120)
+
+    y5 = 1 / (1 + np.exp(x/8 - 6))
+    y6 = 1 / (1 + np.exp(x/15 - 6))
+    y7 = 1 / (1 + np.exp(x/25 - 6))
+    y8 = 1 / (1 + np.exp(x/40 - 6))
+    y9 = 1 / (1 + np.exp(x/3 - 6))
+
+    fig, ax = plt.subplots()
+    ax.plot(x, y0, label='y0')
+    ax.plot(x, y1, label='y1')
+    ax.plot(x, y2, label='y2')
+    ax.plot(x, y3, label='y3')
+    ax.plot(x, y4, label='y4')
+
+    #ax.plot(x, y5)
+    #ax.plot(x, y6)
+    #ax.plot(x, y7)
+    #ax.plot(x, y8)
+    #ax.plot(x, y9)
+
+    ax.grid()
+    ax.legend()
+
+    plt.show()
 
 
-x = np.linspace(0, 500, 100)
-y0 = np.exp(-x / 70)
-y1 = np.exp(-x / 5)
-y2 = np.exp(-x / 20)
-y3 = np.exp(-x / 40)
-y4 = np.exp(-x / 120)
+def exp2b():
+    num = 31 #21
+    x = np.logspace(start=-6, stop=-1, num=num, base=10)
+    y = np.linspace(1, 10, num=num)
+    fig, ax = plt.subplots()
+    ax.set_xscale('log')
+    ax.plot(x, y, 'o')
+    ax.grid()
+    for i,j in zip(x,y):
+        ax.annotate(str(i),xy=(i,j))
+    plt.show()
+    print(x)
 
-y5 = 1 / (1 + np.exp(x/8 - 6))
-y6 = 1 / (1 + np.exp(x/15 - 6))
-y7 = 1 / (1 + np.exp(x/25 - 6))
[…]

### 2019-06-25 project_2/generate_protocol.py
Adding experiment 5 (a, b, c)

@@ -1,0 +1,35 @@
+import numpy as np
+import json
+import os
+
+
+def generate_json(experiment, num_alphas=31):
+    protocol = {
+        'experiments': [
+
+        ],
+        'global_params': {
+            'optimizer': 'rmsprop',
+            'batch_size': 32,
+            'layers': [512, 512],
+            'max_episodes': 1000,
+            'stop_when_solved': False,
+            'keep_learning': True,
+            'print_frequency': 50
+        }
+    }
+    a = np.logspace(start=-6, stop=-1, num=num_alphas, base=10)
+    for i in range(num_alphas):
+        e = {"id": f'E{i + 1:02d}', "dueling": True, "double": True,
+             "prioritized_er": True, "alpha": a[i], "gamma": 0.99}
+        protocol['experiments'].append(e)
+
+    if not os.path.exists(f'experiments/{experiment}'):
+        os.makedirs(f'experiments/{experiment}')
+
+    with open(f'experiments/{experiment}/protocol.json', 'w') as fp:
+        json.dump(protocol, fp, indent=4)
+
+
+if __name__ == '__main__':
+    generate_json(experiment='e5a')

### 2019-06-28 project_2/playground.py
Finishing paper introduction and methods, and adding more experiments final results

@@ -23,6 +23,8 @@ def exp_curves():
     ax.plot(x, y2, label='y2')
     ax.plot(x, y3, label='y3')
     ax.plot(x, y4, label='y4')
+    ax.axhline(0.05, label='base')
+
 
     #ax.plot(x, y5)
     #ax.plot(x, y6)
@@ -51,4 +53,4 @@ def exp2b():
 
 
 if __name__ == '__main__':
-    exp2b()
+    exp_curves()

### 2019-07-01 project_2/paper/paper.tex
Adding final results and final report

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}		% Can be removed after putting your text content
\usepackage{textgreek}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{algpascal}
\usepackage{graphicx}
\usepackage[
backend=biber,
style=numeric,
sorting=ynt,
citestyle=authoryear
]{biblatex}
\addbibresource{paper.bib}

%\graphicspath{ {./images/} }

\title{Lunar Lander with Deep Reinforcement Learning}

\date{June 30th, 2019}

\author{
Carlos Souza\thanks{Latest Git hash: INCLUDE LATEST GITHASH}\\
%Department of Computer Science\\
Georgia Institute of Technology\\
%São Paulo, SP, Brazil \\
\texttt{souza@gatech.edu} \\
}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\begin{document}
    \maketitle

    % keywords can be removed
    %\keywords{Reinforcement learning \and Deep Reinforcement Learning \and Deep Q-Networks}
    %\githash{DO NOT FORGET TO INCLUDE LATEST GIT HASH}

    \section{Introduction}
    \label{sec:introduction}
    Reinforcement learning (RL) theory provides mathematical background and classic algorithms to enable agents to learn optimally control an environment.
    However, to successfully apply those algorithms in problems with real-world complexity, several challenges must be tackled, must notably i) how to derive efficient representations of the environment from high-dimensional sensory inputs, and ii) how to use these to generalize past experience to new situation (\cite{dqn}).

    Recent advances in deep neural networks (DNNs) originated a new type of agent, known as Deep Q-Network agent, capable of overcoming these challenges, which enabled RL algorithms to perform effectively.
    All the many recent successes in applying RL to complex sequential decision-making problems were kick-started by the Deep Q-Network algorithm (DQN; \cite{dqn}).
    Since then, many extensions have been proposed to improve its stability and/or speed.
Double DQN (DDQN; \cite{ddqn}) addresses the problem of overestimation of action values, by decomposing the max operation in the target into action selection and action evaluation.
    The Dueling Network architecture (Dueling DDQN; \cite{dueling}) better generalize learning accross actions by proposing an architecture that explicitly separates the representation of state values and (state-dependent) action advantages.
    Prioritized Experience Replay (PER; \cite{per}) improves data efficiency, by replaying more often transitions from which there is more to learn.

    Each of these algorithms enables substantial performance improvements in isolation and combined, since they build on a shared framework.
    In this paper we propose to study an agent that combines all the aforementioned improvements, exploring its performance in Lunar Lander environment from OpenAI gym (\cite{openai}).

    \subsection{Background}
    \label{subsec:background}
    One of the early breakthroughs in RL was the development of an off-policy Temporal Difference control algorithm known as \emph{Q-learning} (\cite{qlearning}), defined by:
    \begin{equation}
        Q(s_{t}, a_{t}) \leftarrow Q(s_{t}, a_{t}) + \alpha \left[ r_{t+1} + \gamma \max_{a} Q(s_{t+1}, a) - Q(s_{t}, a_{t}) \right],
    \end{equation}
    where $\alpha$ is a constant step-size parameter, or learning rate.
    The learned action-value function $Q$ directly approximates the \emph{optimal} action-value function $Q^{*}$, independent of the policy being followed.

    However, large state and/or action spaces make it intractable to learn $Q$ value estimates for each state-action pairs independently.
    To solve this challenge, DQN (\cite{dqn}) successfully used deep neural networks to approximate the state-action value function as $Q(s, a; \theta)$, where $\theta$ are the parameters of the network.
    For an $n$-dimensional state space and an action space containing $m$ actions, the neural network is a function from $\mathbb{R}^{n}$ to $\mathbb{R}^{m}$.

    Two important components of the DQN algorithm proposed by \cite{dqn} are i) the use of a target network, and ii) the use of experience replay.
    At each time step, the agent selects an action $\epsilon$-greedily with respect to the action values, and adds a transition $\langle s_{t}, a_{t}, r_{t}, s_{t+1} \rangle$ to the experience replay memory, that holds millions of experience tuples.
    The parameters of the neural network are then optimized by using stochastic gradient descent to minimize the loss function at iteration $i$:
    \begin{align}
        L_{i}(\theta_{i}) &= \mathbb{E}_{s, a, r, s'} \left[ \left( y_{i} - Q(s, a; \theta_{i}) \right)^2 \right] \\
        y_{i} &= r + \gamma \max_{a'}Q(s', a'; \theta^{-})
    \end{align}
    where $\theta^{-}$ are the parameters of a fixed and separate \emph{target network}.
    The parameters of the target network $Q(s', a'; \theta^{-})$ are frozen for a fixed number of iterations while updating the \emph{online network} $Q(s, a; \theta_{i})$ by gradient descent.
    The optimization is performed on mini-batches sampled uniformly at random from the experience replay memory.
    These two additions greatly improves the stability of the algorithm, leading to super-human performance on several Atari games.

    \subsection{DQN Extensions}
    \label{subsec:extensions}

    \paragraph{Double Q-learning.}
    The max operator in standard Q-learning and DQN uses the same values both to select and to evaluate an action.
    This makes it more likely to select overestimated values, resulting in overoptimistic value estimates.
    To prevent this, we can decouple the selection from the evaluation.
    This is the idea behind Double Q-learning.
    It is possible to effectively combine this with DQN using the loss
    \begin{align}
        L_{i}(\theta_{i}) &= \mathbb{E}_{s, a, r, s'} \left[ \left( y_{i} - Q(s, a; \theta_{i}) \right)^2 \right] \\
        y_{i} &= r + \gamma Q(s', \argmax_{a'} Q(s', a'; \theta_{i}); \theta^{-}),
    \end{align}
    which was shown to improve performance by reducing value overestimates present in vanilla DQN.

    \paragraph{Dueling networks.}
    The dueling network is a DNN architecture designed for value based RL.
    The key insight behind it is that, for many states, it is unnecessary to estimate the value of each action choice.
    After the first DQN original layers, two streams of fully-connected layers are constructed such that they have the capability of providing separate estimates of the value and advantage functions, being the \emph{advantage function} defined as $A^{\pi}(s, a) = Q^{\pi}(s, a) - V^{\pi}(s)$.
    Finally, the two streams are combined to produce a single output $Q$ function
    \begin{equation}
        Q(s, a; \theta, \alpha, \beta) = V(s; \theta, \beta) + \left( A(s, a; \theta, \alpha) - \frac{1}{|\mathcal{A}|} \sum_{a'} A(s, a'; \theta, \alpha) \right),
    \end{equation}
    where $\theta$ denotes the parameters of the previous layers, while $\alpha$ and $\beta$ are the parameters of the two streams of fully-connected layers.
    As the dueling architecture shares the same input-output interface with standard Q networks, we can recycle all learning algorithms with Q networks to train the dueling architecture.

    \paragraph{Prioritized experience replay.}
    DQN samples uniformly from the replay memory.
    Their key idea behind prioritized experience replay is to increase the replay probability of experience tuples from which there is much to learn.
    \cite{per} proposes the last encountered absolute \emph{TD error} as a proxy for learning potential, which is the central component of the algorithm.

    Several implementation challenges arise from the greedy TD-error prioritization and the bias it introduces by changing the distribution of experience samples.
    To overcome those, the algorithm implements \emph{stochastic prioritization} and \emph{weighted importance-sampling}.

    Prioritized experience replay is shown to deliver both faster learning and better final policy quality across most games of the Atari benchmark suite, as compared to uniform experience replay.

    \section{Methods}
    \label{sec:methods}
    To investigate how the integrated agent that combines all described improvements perform, using Lunar Lander environment as test-bed, we propose 5 different experiments detailed below.

    \paragraph{Experiment 1: Impact of different learning protocols and DQN extensions.}
    We started our study by analyzing the impact of two learning protocols in 8 different agents.
    In the first learning protocol, agents kept learning for 1000 episodes.
    In the second, we stopped the learning as soon as 100-episodes rolling average of the score hit 200, acting greedily from that point on until reaching 1000 episodes, for comparison.
    The 8 different agents combine all DQN extensions, representing all possible combinations.
    The simplest agent implements vanilla DQN, while the most sophisticated, the integrated agent, combines all extensions described earlier.
    We also compared these with agents containing only one and/or two extensions, to understand the impact of each extension on their performance.

    While varying the extensions, we kept all other hyper-parameters frozen.
    We used RMSProp as the optimization algorithm, with $\alpha$ (learning rate) of $0.0001$.
    $\gamma$ (discount factor) was set at $0.99$.
    Replay memory capacity was set to $10,000$ experience tuples, and training was performed using $32$ as mini-batch size.
    The network architecture consists of 2 fully-connected hidden layers with 512 neurons each, both with rectified linear unit (ReLU) activation functions (for dueling network: we added a fully-connected layer as described previously).
    For prioritized replay, we annealed the bias by progressively increasing $\beta$ (exponent in importance-sampling weights) from $0.4$ to $1$.
    As exploration strategy, we decayed $\epsilon$ following $\epsilon(t) = e^{-t/70}$, where $t$ is the current episode, until hitting $0.05$, staying constant at $0.05$ from that point on.

    To reduce the variability given the stochasticity, this experiment was performed 20 times for each variant.

    \paragraph{Experiment 2: Effect of different gradient descent optimization algorithms and learning rates.}
    Next, we assessed the effect of different optimization algorithms and their optimal learning rates on the integrated agent, which combines all DQN extensions.
    We simulated RMSProp, Adaptive Moment Estimation (Adam), and Stochastic Gradient Descent (SGD).
    For those 3 different algorithms, we tested $30$ different $\alpha$ (learning rates), evenly spaced on a log scale from $10^{-6}$ to $10^{-1}$.
    All other hyper-parameters were kept frozen as described in experiment 1.

    \paragraph{Experiment 3: Discount factor sensitivity.}
    In this experiment, we investigated the effect of $\gamma$ in the integrated agent's performance.
    We tested all $\gamma \in [0.8, 0.85, 0.9, 0.92, 0.94, 0.96, 0.98, 0.99, 1.0]$.
    As usual, all other hyper-parameters were kept frozen.
    To reduce the variability given the stochasticity, this experiment was performed 5 times for each variant.

    \paragraph{Experiment 4: Evaluation of different exploration strategies.}
    Following, we evaluated the impact of $10$ different $\epsilon$ decay curves.
    We assessed two types of curves: exponential decay functions, $\epsilon(t) = e^{-k t}$; and inverted logistic functions, $\epsilon(t) = 1 / (1 + e^{-k_{1} (t - k_{2})})$.
    Again, all other hyper-parameters were kept frozen.
    To reduce the variability given the stochasticity, this experiment was performed 10 times for each variant.

    \section{Results}
    \label{sec:results}

    \begin{figure}[t]
        \includegraphics[width=\textwidth]{./images/fig1.png}
        \centering
        \caption{Different learning protocols.
        In the left, integrated agent (combining all DQN extensions) learns continuously for 1000 episodes.
        In the center, integrated agent learns just until it achieves 200 score over 100-episodes rolling average, acting greedily from that point on until 1000 episodes.
        In the right, we test both learning protocols, comparing both trained agents (no learning).
        All results presented are averages of 20 runs to reduce the variance: the shaded area is the standard deviation.}
        \label{fig:fig1}
    \end{figure}

    Figure \ref{fig:fig1} presents first part of experiment 1, in which we confirmed our expectations.
    In both learning protocols, the integrated agent, combining all DQN extensions, solved Lunar Lander environment by reaching 200 score over past 100-episodes rolling average around the same time, in 230 and 226 episodes respectively.
    Also, as expected, the learning protocol that kept the agent learning for the whole 1000 episodes achieved a higher score in the 100-episodes test, 259 in average, in comparison to the learning protocol that stopped the learning after solving the environment (only 237 in average).

    \begin{figure}[h]
        \includegraphics[width=\textwidth]{./images/fig2.png}
        \centering
        \caption{Different agents.
        The 2 charts in the left show the first learning protocol, where agents keep learning for 1000 episodes, while the 2 charts in the right show the second learning protocol, where agents stop learning right after reaching 200 score on 100-episodes rolling average.
        To better visualize the test score results, comparing the performance of all different agents over 100 episodes with no learning, in second and fourth chart we used boxplots, for the first and second learning protocols respectively.}
        \label{fig:fig2}
    \end{figure}

    Figure \ref{fig:fig2} presents second part of experiment 1.
    First 2 charts show first learning protocol (all agents keep learning for 1000 episodes), while last 2 charts show second learning protocol (agents stop learning as soon as they hit 200 score).
    Regardless of the learning protocol, all agents solve Lunar Lander environment in 200 to 250 episodes.

    In the first learning protocol, the agent with the second worst performance was the vanilla DQN, with 251 score over 100-episodes test, contradicting our expectation: we expected this agent to show the poorest performance.
    The agent with the worst performance implemented DQN with Dueling Networks and Prioritized Experience Replay (DuelDQN+PER).
    Also contradicting our expectation, two agents showed better performance than the integrated agent (257 score): the agent with Dueling Networks + Double DQN (DuelDDQN, with 263 score), and the DQN agent with Prioritized Experience Replay (DQN+PER, with 267 score).

    In the second learning protocol, the best agent was the integrated agent (237 score), which combines all DQN extensions, confirming our expectation.
    However, contradicting our expectation, in this case the worst performance was not the vanilla DQN, but surprisingly the DQN agent with Prioritized Experience Replay (PER, with 204 score), which had performed best in the first learning protocol.

    These results suggest that the integrated agent is the best choice in limited/short training time, as it scored the highest amongst all options when learning happened only until 100-episodes rolling average hit 200 score.
    They also suggest that the Prioritized Experience Replay extension benefit is only perceived when learning happens for longer.
    Another interesting observation is regarding the agent with Double DQN and Prioritized Experience Replay (DDQN+PER).
    It performed second worse in first learning protocol, but second best in second learning protocol, which suggests those extensions, combined, are great tools to improve the performance in limited training time, but are not as interesting when we have time to train for longer periods.

    \begin{figure}[t]
        \includegraphics[width=\textwidth]{./images/fig3.png}
        \centering
        \caption{First 3 charts show best score achieved by integrated agent over 1000 episodes using different optimization algorithms and learning rates, while chart in the right shows score evolution over 1000 episodes for different discount factors.}
        \label{fig:fig3}
    \end{figure}

    Figure \ref{fig:fig3} presents experiment 2 and 3.
    All optimization algorithms produced comparable best performances: RMSProp best result was 282 score for $\alpha = 6.8 \times 10^{-5}$, Adaptive Moment Estimation (Adam) best result was 281 score for $\alpha = 3.2 \times 10^{-4}$, and Stochastic Gradient Descent (SGD) best result was 277 for $\alpha = 10^{-2}$.

    An interesting observation can be made regarding how many tested learning rates provided scores above 200.
    13 out of 30 tested learning rates achieved results above 200 score for Adam algorithm;
    11 in the case of RMSProp;
    and only 6 in the case of SGD.
    This broader range of learning rates with superior performance for Adam and RMSProp in comparison to SGD suggests that those two algorithms should be preferred over SGD, which confirms our expectation.

    In figure \ref{fig:fig3} we can also see the impact of discount factor in agent's convergence.
    From the different discount factors tested, only $\gamma = 0.98$ and $\gamma = 0.99$ enabled the integrated agent to solve the Lunar Lander environment, converging to scores above 200.

    \begin{figure}[h]
        \includegraphics[width=\textwidth]{./images/fig4.png}
        \centering
        \caption{Impact of different exploration strategies in agent's learning.
        In the left, we show 10 different $\epsilon$-decay curves used, while in the right we show the effect of each of these curves in the integrated agent's performance.}
        \label{fig:fig4}
    \end{figure}

    Results from our final experiment are present in figure \ref{fig:fig4}.
    All exploration strategies enabled the integrated agent to converge and solve Lunar Lander environment regardless of their $\epsilon$-decay curve format, as expected.
    However, some exploration strategies converged much faster.
    Curve 4, described by $\epsilon(t) = e^{-t / 40}$, showed the best result, making the agent achieve 200 score over 100-episodes rolling average in 219 episodes.
    Curve 10, described by $\epsilon(t) = 1 / (1 + e^{t/40 - 6})$, showed the worst result, converging in 390 episodes.
    Curiously, curves 2, 3 and 6, which decay $\epsilon$ more aggressively than curve 4, performed worse than the best curve.
    This observation is yet another occurrence of something noted in several experiments: the best performing hyper-parameter is never at extremes.

    \section{Conclusion}
    \label{sec:conclusion}

    In this work, we implemented Deep Q-Networks and some of their main extensions (Double Q-learning, Dueling Networks, and Prioritized Experience Replay), studying their performance in Lunar Lander environment from OpenAI gym.
    We were able to confirm several points made in the articles that introduced DQN and their extensions.
    Our experiments achieved results that would put our implemented agents in top 2nd position in the \href{https://github.com/openai/gym/wiki/Leaderboard#lunarlander-v2}{leaderboard}.

\printbibliography

\end{document}

### 2019-07-01 project_2/agent_pytorch.py

@@ -152,37 +152,40 @@ class Agent(object):
             else:
                 batch = random.sample(self.memory, self.batch_size)
 
-            states = torch.stack([torch.from_numpy(i[0]) for i in batch]).float().to(self.device)
-            y = self.Q(states).detach().cpu().numpy()
-            next_states = torch.stack([torch.from_numpy(i[3]) for i in batch]).float().to(self.device)
-            q_t = self.Q_target(next_states).detach().cpu().numpy()
+            try:
+                states = torch.stack([torch.from_numpy(i[0]) for i in batch]).float().to(self.device)
+                y = self.Q(states).detach().cpu().numpy()
+                next_states = torch.stack([torch.from_numpy(i[3]) for i in batch]).float().to(self.device)
+                q_t = self.Q_target(next_states).detach().cpu().numpy()
 
-            # Do Q-learning update
-            for i, (s, a, r, s_p, d) in enumerate(batch):
-                if d:
-                    y[i][a] = r
-                else:
-                    y[i][a] = r + self.gamma * np.max(q_t[i])
+                # Do Q-learning update
+                for i, (s, a, r, s_p, d) in enumerate(batch):
+                    if d:
+                        y[i][a] = r
+                    else:
+                        y[i][a] = r + self.gamma * np.max(q_t[i])
 
-            # Perform a gradient descent step
-            target = torch.from_numpy(y).to(self.device)
-            old = self.Q(states).to(self.device)
-            if self.prioritized_er:
-                if self.huber:
-                    t = torch.abs(target - old).to(self.device)
-                    loss = (is_weights * torch.where(t < 1, 0.5 * t ** 2, t - 0.5)).mean().to(self.device)
+                # Perform a gradient descent step
+                target = torch.from_numpy(y).to(self.device)
+                old = self.Q(states).to(self.device)
+                if self.prioritized_er:
+                    if self.huber:
+                        t = torch.abs(target - old).to(self.device)
+                        loss = (is_weights * torch.where(t < 1, 0.5 * t ** 2, t - 0.5)).mean().to(self.device)
+                    else:
+                        loss = (is_weights * ((target - old) ** 2)).mean().to(self.device)
                 else:
-                    loss = (is_weights * ((target - old) ** 2)).mean().to(self.device)
-            else:
-                if self.huber:
-                    t = torch.abs(target - old).to(self.device)
-                    loss = torch.where(t < 1, 0.5 * t ** 2, t - 0.5).to(self.device)
-                else:
[…]

### 2019-07-01 project_2/paper/paper.tex
Fixing typo in README.md

@@
     \label{sec:methods}
-    To investigate how the integrated agent that combines all described improvements perform, using Lunar Lander environment as test-bed, we propose 5 different experiments detailed below.
+    To investigate how the integrated agent that combines all described improvements perform, using Lunar Lander environment as test-bed, we propose 4 different experiments detailed below.
 

### 2019-07-01 project_2/paper/paper.tex
Adding final paper

@@
 \author{
-Carlos Souza\thanks{Latest Git hash: INCLUDE LATEST GITHASH}\\
+Carlos Souza\thanks{Latest Git hash: 5b33898a39727cea2d7f0a0872acdd30f9d2a1e5}\\
 %Department of Computer Science\\
 Georgia Institute of Technology\\
-%São Paulo, SP, Brazil \\
+São Paulo, SP, Brazil \\
 \texttt{souza@gatech.edu} \\
@@
     As exploration strategy, we decayed $\epsilon$ following $\epsilon(t) = e^{-t/70}$, where $t$ is the current episode, until hitting $0.05$, staying constant at $0.05$ from that point on.
-
-    To reduce the variability given the stochasticity, this experiment was performed 20 times for each variant.
+    To reduce the variance given the stochasticity, this experiment was performed 20 times for each variant.
 
@@
     As usual, all other hyper-parameters were kept frozen.
-    To reduce the variability given the stochasticity, this experiment was performed 5 times for each variant.
+    To reduce the variance given the stochasticity, this experiment was performed 5 times for each variant.
 
@@
     Again, all other hyper-parameters were kept frozen.
-    To reduce the variability given the stochasticity, this experiment was performed 10 times for each variant.
+    To reduce the variance given the stochasticity, this experiment was performed 10 times for each variant.
 

### 2019-07-11 project_2/paper/deleted.tex
Adding HW5 and HW6

In the standard reinforcement learning setting, agent interacts with an environment over discrete time steps.
    At each time step $t$, the agent perceives a state $s_{t} \in \mathcal{S}$, chooses an action from a discrete set $a_{t} \in \mathcal{A}$ accordingly to a policy $\pi$, where $\pi$ is a mapping from state space $\mathcal{S}$ to action space $\mathcal{A}$, and observes a reward $r_{t}$ and a next state $s_{t+1}$.
    This process continues until the agent reaches a terminal state, after which the process is restarted.
    The purpose of the agent is to maximize the expected discounted reward $R_{t} = \sum_{k=1}^\infty \gamma^k r_{k+t}$, where $\gamma \in (0, 1]$ is the discount factor that trades-off the importance of immediate vs future rewards.
    This interaction between the agent and the environment is formalized as a \emph{Markov Decision Process} (MDP), described by $\langle \mathcal{S}, \mathcal{A}, T, r, \gamma \rangle$ tuple, where $T(s, a, s') = P[s_{t+1} = s' | s_{t} = s, a_{t} = a]$ is the stochastic transition function.

For an agent following a stochastic policy $\pi$, the true value of an action $a$ in a state $s$, and the value of that state $s$ are:
    \begin{align}
        Q^{\pi}(s, a) &= \mathbb{E}[R_{t} | s_{t} = s, a_{t} = a], \text{and} \label{eq:1} \\
        V^{\pi}(s) &= \mathbb{E}_{a\sim \pi(s)}[Q^{\pi}(s, a)] \label{eq:2}
    \end{align}
    The optimal state-value function is defined as $Q^{*}(s, a) = \max_{\pi}Q^{\pi}(s, a)$.
    Under deterministic policy $a = \argmax_{a' \in \mathcal{A}}Q^{*}(s, a')$, it follows that $V^{*} = \max_{a}Q^{*}(s, a)$.
    Also, the optimal state-value function $Q$ satisfies the Bellman equation:
    \begin{equation}
        Q^{*}(s, a) = \mathbb{E}_{s'}\left[r + \gamma \max_{a'} Q^{*}(s', a') \:|\: s, a\right]
    \end{equation}

    A common way of deriving a new policy from a state-action value function is to act $\epsilon$-greedly with respect to the action values, i.e. taking the action with the highest value with probability $(1 - \epsilon)$ or otherwise acting randomly with probability $\epsilon$.
    The optimal policy is easily derived from the optimal state-action value function by selecting the highest-valued action in each state.

\subsection{Lunar Lander Environment}
    \label{subsec:lunarlander}
    Rocket trajectory optimization is a classic topic in Optimal Control.
    OpenAI Gym proposes an environment where an agent has to control a "Lunar Lander" with the purpose of landing it successfully in a landing pad.
    The agent has four possible actions: do nothing, fire the left orientation engine, fire the main engine, or fire the right orientation engine.
    The state space is a 8-dimensional continuous space: at each time step, environment returns Lunar Lander position (x, y), velocities (horizontal and vertical), angle, angular velocity, and whether left/right legs are in contact with the ground.
    Landing pad coordinates are (0,0).
    Reward for moving from the top of the screen to landing pad and zero speed is about 100..140 points.
    If lander moves away from landing pad it loses reward back.
    Episode finishes if the lander crashes or comes to rest, receiving additional -100 or +100 points.
    Each leg ground contact is +10.
    Firing main engine is -0.3 points each frame.
    Firing side engine is -0.03 points each frame.
    The problem is considered to be solved when the agent achieves 200 score over past 100-episodes rolling average.

    \paragraph{Experiment 5: Effect of kick-starting replay memory with human experience.}
    Finally, we investigate how kick-starting the replay memory with human experience can speed-up agent learning.
    To do it, we generate experience tuples $\langle s, a, s', r \rangle$ by playing Lunar Lander by hand, and use this data to pre-populate the replay memory.
    Then, we set our integrated agent to start learning procedure, with $\epsilon$ constant at $0.05$.

### 2019-07-11 gym/examples/agents/keyboard_agent.py

@@ -1,0 +1,101 @@
+#!/usr/bin/env python
+from __future__ import print_function
+
+import sys, gym, time
+
+#
+# Test yourself as a learning agent! Pass environment name as a command-line argument, for example:
+#
+# python keyboard_agent.py SpaceInvadersNoFrameskip-v4
+#
+
+env = gym.make('LunarLander-v2' if len(sys.argv)<2 else sys.argv[1])
+
+if not hasattr(env.action_space, 'n'):
+    raise Exception('Keyboard agent only supports discrete action spaces')
+ACTIONS = env.action_space.n
+SKIP_CONTROL = 0    # Use previous control decision SKIP_CONTROL times, that's how you
+                    # can test what skip is still usable.
+
+human_agent_action = 0
+human_wants_restart = False
+human_sets_pause = False
+
+
+def key_to_action(key):
+    if key == 65363:
+        return 1
+    elif key == 65362:
+        return 2
+    elif key == 65361:
+        return 3
+    else:
+        return -1
+
+def key_press(key, mod):
+    print(key)
+    global human_agent_action, human_wants_restart, human_sets_pause
+    if key==0xff0d: human_wants_restart = True
+    if key==32: human_sets_pause = not human_sets_pause
+    a = key_to_action(key)
+    #a = int( key - ord('0') )
+    if a <= 0 or a >= ACTIONS: return
+    human_agent_action = a
+
+
+def key_release(key, mod):
+    global human_agent_action
+    a = key_to_action(key)
+    #a = int( key - ord('0') )
+    if a <= 0 or a >= ACTIONS: return
+    if human_agent_action == a:
+        human_agent_action = 0
+
+
+env.render()
+env.unwrapped.viewer.window.on_key_press = key_press
+env.unwrapped.viewer.window.on_key_release = key_release
+
+
+def rollout(env):
+    global human_agent_action, human_wants_restart, human_sets_pause
+    human_wants_restart = False
+    state = env.reset()
+    skip = 0
+    total_reward = 0
+    total_timesteps = 0
+    while 1:
+        if not skip:
+            #print("taking action {}".format(human_agent_action))
+            a = human_agent_action
+            total_timesteps += 1
+            skip = SKIP_CONTROL
+        else:
+            skip -= 1
+
+        next_state, r, done, info = env.step(a)
+        #if r != 0:
+        #    print("reward %0.3f" % r)
+        total_reward += r
+
+        state = next_state
+
[…]

### 2019-07-11 gym/setup.py

@@ -1,0 +1,58 @@
+from setuptools import setup, find_packages
+import sys, os.path
+
+# Don't import gym module here, since deps may not be installed
+sys.path.insert(0, os.path.join(os.path.dirname(__file__), 'gym'))
+from version import VERSION
+
+# Environment-specific dependencies.
+extras = {
+  'atari': ['atari_py~=0.2.0', 'Pillow', 'opencv-python'],
+  'box2d': ['box2d-py~=2.3.5'],
+  'classic_control': [],
+  'mujoco': ['mujoco_py>=1.50, <2.1', 'imageio'],
+  'robotics': ['mujoco_py>=1.50, <2.1', 'imageio'],
+}
+
+# Meta dependency groups.
+all_deps = []
+for group_name in extras:
+    all_deps += extras[group_name]
+extras['all'] = all_deps
+
+setup(name='gym',
+      version=VERSION,
+      description='The OpenAI Gym: A toolkit for developing and comparing your reinforcement learning agents.',
+      url='https://github.com/openai/gym',
+      author='OpenAI',
+      author_email='gym@openai.com',
+      license='',
+      packages=[package for package in find_packages()
+                if package.startswith('gym')],
+      zip_safe=False,
+      install_requires=[
+          'scipy', 'numpy>=1.10.4', 'six', 'pyglet>=1.2.0', 'cloudpickle~=1.2.0',
+          'enum34~=1.1.6;python_version<"3.4"'
+      ],
+      extras_require=extras,
+      package_data={'gym': [
+        'envs/mujoco/assets/*.xml',
+        'envs/classic_control/assets/*.png',
+        'envs/robotics/assets/LICENSE.md',
+        'envs/robotics/assets/fetch/*.xml',
+        'envs/robotics/assets/hand/*.xml',
+        'envs/robotics/assets/stls/fetch/*.stl',
+        'envs/robotics/assets/stls/hand/*.stl',
+        'envs/robotics/assets/textures/*.png']
+      },
+      tests_require=['pytest', 'mock'],
+      python_requires='>=2.7, !=3.0.*, !=3.1.*, !=3.2.*, !=3.3.*, !=3.4.*',
+      classifiers=[
+          'Programming Language :: Python :: 2',
+          'Programming Language :: Python :: 2.7',
+          'Programming Language :: Python :: 3',
+          'Programming Language :: Python :: 3.5',
+          'Programming Language :: Python :: 3.6',
+          'Programming Language :: Python :: 3.7',
+      ],
+)

### 2019-07-11 homeworks/hw5/main.py

@@ -1,0 +1,91 @@
+import numpy as np
+
+
+class KWIKSolver(object):
+    def __init__(self, X, Y):
+        self.X = X
+        self.Y = Y
+        self.num_episodes = len(Y)
+        self.hypothesis = np.full(len(X[0]), -1)
+        self.memory = {'x': [], 'y': []}
+
+    def update_hypothesis(self, x, y):
+        if y == 1:
+            for i in range(len(x)):
+                if x[i] == 1:
+                    self.hypothesis[i] = 1
+
+        elif y == 0:
+            pass
+
+    def update_memory(self, x, y):
+        self.memory['x'].append(x)
+        self.memory['y'].append(y)
+
+    def retrieve_from_memory(self, x):
+        for i in range(len(self.memory['x'])):
+            if x == self.memory['x'][i]:
+                return self.memory['y'][i]
+
+        return None
+
+    def predict(self, x):
+        # First check whether there is a prediction in memory
+        prediction = self.retrieve_from_memory(x)
+        if prediction is not None:
+            print('from memory')
+            return prediction
+
+        # Get the hypothesis for the people present
+        hyp_present = np.ma.masked_array(self.hypothesis,
+                                         mask=1 - np.array(x)).compressed()
+
+        # If there's no one, no fight
+        if len(hyp_present) == 0:
+            print('no one')
+            return 0
+
+        # If uncertainty
+        if np.all(hyp_present == -1):
+            print('uncertainty')
+            return -1
+
+        # If instigator might be present, and there's no chance of a peacemaker
+        if np.any(hyp_present == 1) and not np.any(hyp_present == 0):
+            print('instigator')
+            return 1
+
+        # If there's a chance of a peacemaker
+        if np.any(hyp_present == 0):
+            print('peacemaker')
+            return 0
+
+    def solve(self):
+        result = [self.Y[0]]
+        self.update_memory(self.X[0], self.Y[0])
+        for i in range(1, self.num_episodes):
+            self.update_hypothesis(self.X[i - 1], self.Y[i - 1])
+            prediction = self.predict(self.X[i])
+            result.append(prediction)
+            if prediction != -1:
+                self.update_memory(self.X[i], prediction)
+
+            print(f'x: {self.X[i]}\npred: {prediction}\ny: {self.Y[i]}\nhypothesis: {self.hypothesis}\n')
+
+        return result
+
[…]

### 2019-07-11 homeworks/hw5/main2.py

@@ -1,0 +1,73 @@
+import re
+import json
+from itertools import permutations
+
+
+def pred_or_learn(patrons, num_patrons, H, fight):
+
+    # if all patrons present, no fight will occur for sure
+    if sum(patrons) == num_patrons:
+        return (H, 0)
+
+    # store the opinions from each hypothesis h
+    votes = []
+
+    # itearte each h in Hypothesis space H
+    for h in H:
+        # from h, compute logics whether instigator/peacemaker present in the current episode
+        instigator_ind, peacemaker_ind = h[0], h[1]
+        instigator_presence = (patrons[instigator_ind] == 1)
+        peacemaker_presence = (patrons[peacemaker_ind] == 1)
+
+        # fight only occurs if instigator presents while peacemake does not
+        if (instigator_presence and not peacemaker_presence):
+            votes.append(1)
+        else:
+            votes.append(0)
+
+    # check whether all remaining h have unaminous opinions on fight or no-fight
+    # if Yes |L| = 1, the algo knows the proper output
+    if (sum(votes) == len(votes) or sum(votes) == 0):
+        return (H, votes[0])
+
[…]

### 2019-07-11 homeworks/hw6/main.py

@@ -1,0 +1,73 @@
+from pulp import *
+import numpy as np
+
+
+def test_pulp():
+    prob = LpProblem("The Whiskas Problem",LpMinimize)
+
+    # The 2 variables Beef and Chicken are created with a lower limit of zero
+    x1=LpVariable("ChickenPercent",0,None,LpInteger)
+    x2=LpVariable("BeefPercent",0)
+
+    # The objective function is added to 'prob' first
+    prob += 0.013*x1 + 0.008*x2, "Total Cost of Ingredients per can"
+
+    # The five constraints are entered
+    prob += x1 + x2 == 100, "PercentagesSum"
+    prob += 0.100*x1 + 0.200*x2 >= 8.0, "ProteinRequirement"
+    prob += 0.080*x1 + 0.100*x2 >= 6.0, "FatRequirement"
+    prob += 0.001*x1 + 0.005*x2 <= 2.0, "FibreRequirement"
+    prob += 0.002*x1 + 0.005*x2 <= 0.4, "SaltRequirement"
+
+    # The problem data is written to an .lp file
+    prob.writeLP("WhiskasModel.lp")
+
+    # The problem is solved using PuLP's choice of Solver
+    prob.solve()
+
+    # The status of the solution is printed to the screen
+    print("Status:", LpStatus[prob.status])
+
+    # Each of the variables is printed with it's resolved optimum value
+    for v in prob.variables():
+        print(v.name, "=", v.varValue)
+
+    # The optimised objective function value is printed to the screen
+    print("Total Cost of Ingredients per can = ", value(prob.objective))
+
+
+def solve_rps(r):
+    prob = LpProblem("Rock, Paper, Scissor", LpMaximize)
+
+    v = LpVariable("TotalValue")
+    p_r = LpVariable("RockProbability", 0, 1)
+    p_p = LpVariable("PaperProbability", 0, 1)
+    p_s = LpVariable("ScissorProbability", 0, 1)
+
+    prob += v
+
+    prob += p_r + p_p + p_s == 1
+    prob += r[0][0] * p_r + r[1][0] * p_p + r[2][0] * p_s >= v
+    prob += r[0][1] * p_r + r[1][1] * p_p + r[2][1] * p_s >= v
+    prob += r[0][2] * p_r + r[1][2] * p_p + r[2][2] * p_s >= v
+
+    prob.writeLP("rps.lp")
+    prob.solve()
+    print("Status:", LpStatus[prob.status])
+    result = [0, 0, 0]
[…]

### 2019-07-12 project_3/environment.py
Adding Soccer environment (needs debugging)

@@ -1,0 +1,196 @@
+# -*- coding: utf-8 -*-
+"""Soccer Environment module.
+
+This module implements Soccer environment as described in Greenwald & Hall's
+`Correlated-Q Learning paper`_. Implementation follows `OpenAI Gym standards`_.
+
+.. _Correlated-Q Learning paper:
+   https://www.aaai.org/Papers/ICML/2003/ICML03-034.pdf
+
+.. _OpenAI Gym standards:
+   http://gym.openai.com/docs/
+
+"""
+
+import numpy as np
+import random
+
+
+class Soccer(object):
+    """Soccer Environment.
+
+    This class implements Soccer environment. All attributes are private.
+    Game simulation should be done using the methods, which return the current
+    state. The current state is a tuple containing:
+    - Position of Player 0 [row, col] in 2x4 grid
+    - Position of Player 1 [row, col] in 2x4 grid
+    - Who is in possession of the ball (0 or 1)
+
+    Attributes:
+        num_actions (int): Number of possible actions in the environment (5).
+
+    """
+    def __init__(self, seed=None):
+        """Initializes the environment.
+
+        Args:
+            seed (int): Random seed. If set, will set Numpy and Random seeds, to
+                enable reproducibility of results.
+        """
+        if seed is not None:
+            np.random.seed(seed)
+            random.seed(seed)
+
+        self.num_actions = 5
+        self._pos_p0 = [None, None]
+        self._pos_p1 = [None, None]
+        self._ball = None
+        self._t = None
+        self.reset()
+
+    def step(self, action_p0, action_p1):
+        """Simulates a step given joint actions from players 0 and 1, returning
+        the new state.
+
+        Args:
+            action_p0 (int): Action from player 0: 1 for North, 2 for East, 3
+                for South, 4 for West, 0 for stay put.
+            action_p1 (int): Action from player 1: 1 for North, 2 for East, 3
+                for South, 4 for West, 0 for stay put.
+
+        Returns:
+            List containing current state components.
+        """
+        self._t += 1
+
+        first_to_act = np.random.randint(2)  # Defines who acts first
+
+        # Updates players positions, and defines ball possession
+        if first_to_act == 0:
+            self._pos_p0 = self._get_new_position(self._pos_p0, action_p0)
[…]

### 2019-07-12 project_3/main.py

@@ -1,0 +1,31 @@
+import numpy as np
+import environment
+
+
+if __name__ == '__main__':
+    env = environment.Soccer()
+    observation = env.reset()
+
+    env._pos_p0 = [1, 1]
+    env._pos_p1 = [1, 2]
+    env._ball = 1
+    env.render()
+
+    observation, rewards, done, info = env.step(3, 4)
+    print(f'Actions: {[3, 4]}, Rewards: {rewards}, Done: {done}, Info: {info}')
+
+    env.render()
+
+    exit(1)
+
+    done = False
+    while not done:
+        env.render()
+        print(observation)
+        a0 = np.random.randint(env.num_actions)
+        a1 = np.random.randint(env.num_actions)
+        observation, rewards, done, info = env.step(a0, a1)
+        print(f'Actions: {[a0, a1]}, Rewards: {rewards}, Done: {done}')
+        if done:
+            env.render()
+            print('DONE!')

### 2019-07-12 project_3/environment.py
Bug fixed

@@ -67,7 +67,10 @@ class Soccer(object):
 
         # Updates players positions, and defines ball possession
         if first_to_act == 0:
-            self._pos_p0 = self._get_new_position(self._pos_p0, action_p0)
+            possible_pos_p0 = self._get_new_position(self._pos_p0, action_p0)
+            if possible_pos_p0 != self._pos_p1:
+                self._pos_p0 = possible_pos_p0
+
             possible_pos_p1 = self._get_new_position(self._pos_p1, action_p1)
             if possible_pos_p1 != self._pos_p0:
                 self._pos_p1 = possible_pos_p1
@@ -75,7 +78,10 @@ class Soccer(object):
                 if self._ball == 1:
                     self._ball = 0
         else:
-            self._pos_p1 = self._get_new_position(self._pos_p1, action_p1)
+            possible_pos_p1 = self._get_new_position(self._pos_p1, action_p1)
+            if possible_pos_p1 != self._pos_p0:
+                self._pos_p1 = possible_pos_p1
+
             possible_pos_p0 = self._get_new_position(self._pos_p0, action_p0)
             if possible_pos_p0 != self._pos_p1:
                 self._pos_p0 = possible_pos_p0

### 2019-07-12 project_3/environment.py
Starting MultiQ solver class

@@ -42,8 +42,8 @@ class Soccer(object):
             random.seed(seed)
 
         self.num_actions = 5
-        self._pos_p0 = [None, None]
-        self._pos_p1 = [None, None]
+        self._pos_p0 = (None, None)
+        self._pos_p1 = (None, None)
         self._ball = None
         self._t = None
         self.reset()
@@ -113,8 +113,8 @@ class Soccer(object):
                 reward_p1 = -100
                 done = True
 
-        return [self._pos_p0, self._pos_p1, self._ball], \
-               [reward_p0, reward_p1], done, [first_to_act, self._t]
+        return (self._pos_p0, self._pos_p1, self._ball), \
+               (reward_p0, reward_p1), done, (first_to_act, self._t)
 
     def _get_new_position(self, curr_pos, action):
         """Updates current position given an action.
@@ -139,13 +139,13 @@ class Soccer(object):
             return curr_pos  # Boundary condition
 
         if action == 1:
-            return [0, curr_pos[1]]
+            return 0, curr_pos[1]
         elif action == 2:
-            return [curr_pos[0], curr_pos[1] + 1]
+            return curr_pos[0], curr_pos[1] + 1
         elif action == 3:
-            return [1, curr_pos[1]]
+            return 1, curr_pos[1]
         elif action == 4:
-            return [curr_pos[0], curr_pos[1] - 1]
+            return curr_pos[0], curr_pos[1] - 1
 
     def reset(self):
         """Resets the environment.
@@ -157,11 +157,11 @@ class Soccer(object):
             List containing current state components.
 
         """
-        self._pos_p0 = [0, 1]
-        self._pos_p1 = [0, 2]
+        self._pos_p0 = (0, 1)
+        self._pos_p1 = (0, 2)
         self._ball = np.random.randint(2)
         self._t = 0
-        return [self._pos_p0, self._pos_p1, self._ball]
+        return self._pos_p0, self._pos_p1, self._ball
 
     def render(self):
         """Renders the environment in screen, for visualization/debugging
@@ -188,15 +188,17 @@ class Soccer(object):
 
 if __name__ == '__main__':
     env = Soccer()
-    observation = env.reset()
+    state = env.reset()
     done = False
     while not done:
         env.render()
-        print(observation)
+        print(f'Player 0 position: {state[0]}, Player 1 position: {state[1]}, '
+              f'Who has the ball: {state[2]}\n')
[…]

### 2019-07-12 project_3/main.py

@@ -1,5 +1,17 @@
 import numpy as np
 import environment
+from collections import defaultdict
+
+
+class MultiQ(object):
+    def __init__(self, selection='Foe-Q', alpha=0.001, gamma=0.9):
+        self.env = environment.Soccer()
+        self.selection = selection
+        self.alpha = alpha
+        self.gamma = gamma
+
+    def train(self):
+        pass
 
 
 if __name__ == '__main__':
@@ -16,16 +28,12 @@ if __name__ == '__main__':
 
     env.render()
 
-    exit(1)
+    Q_matrix = defaultdict(lambda: np.zeros((env.num_actions, env.num_actions)))
+    s_0 = ((0, 1), (1, 2), 0)
+    s_1 = ((1, 1), (0, 0), 1)
 
-    done = False
-    while not done:
-        env.render()
-        print(observation)
-        a0 = np.random.randint(env.num_actions)
-        a1 = np.random.randint(env.num_actions)
-        observation, rewards, done, info = env.step(a0, a1)
-        print(f'Actions: {[a0, a1]}, Rewards: {rewards}, Done: {done}')
-        if done:
-            env.render()
-            print('DONE!')
+    Q_matrix[s_0][1][2] = 1
+    print(Q_matrix[s_0])
+    print(Q_matrix[s_1])
+    Q_matrix[s_0][1][2] += 1
+    print(Q_matrix[s_0])

### 2019-07-12 homeworks/hw6/main.py
Simplifying minimax

@@ -54,20 +54,21 @@ def solve_rps(r):
     prob.writeLP("rps.lp")
     prob.solve()
     print("Status:", LpStatus[prob.status])
-    result = [0, 0, 0]
+
+    list_vars = ['RockProbability', 'PaperProbability', 'ScissorProbability']
+    result = dict()
+    for var in list_vars:
+        result[var] = 0
+
     for v in prob.variables():
         print(v.name, "=", v.varValue)
-        if v.name == 'RockProbability':
-            result[0] = v.varValue
-        elif v.name == 'PaperProbability':
-            result[1] = v.varValue
-        elif v.name == 'ScissorProbability':
-            result[2] = v.varValue
+        if v.name in list_vars:
+            result[v.name] = v.varValue
 
-    return result
+    return list(result.values())
 
 
 if __name__ == '__main__':
-    matrix = np.array([[0.0, 3.79, -1.0], [-3.79, 0.0, 4.98], [1.0, -4.98, 0.0]])
+    matrix = np.array([[0.0, 1.04, -2.07], [-1.04, 0.0, 1.57], [2.07, -1.57, 0.0]])
     result = solve_rps(matrix)
     print(result)

### 2019-07-12 homeworks/hw6/main.py
Simplifying minimax even further

@@ -36,7 +36,7 @@ def test_pulp():
     print("Total Cost of Ingredients per can = ", value(prob.objective))
 
 
-def solve_rps(r):
+def solve_rps(r, verbose=False):
     prob = LpProblem("Rock, Paper, Scissor", LpMaximize)
 
     v = LpVariable("TotalValue")
@@ -45,30 +45,22 @@ def solve_rps(r):
     p_s = LpVariable("ScissorProbability", 0, 1)
 
     prob += v
-
     prob += p_r + p_p + p_s == 1
-    prob += r[0][0] * p_r + r[1][0] * p_p + r[2][0] * p_s >= v
-    prob += r[0][1] * p_r + r[1][1] * p_p + r[2][1] * p_s >= v
-    prob += r[0][2] * p_r + r[1][2] * p_p + r[2][2] * p_s >= v
+    for c in range(r.shape[1]):
+        prob += r[0][c] * p_r + r[1][c] * p_p + r[2][c] * p_s >= v
 
-    prob.writeLP("rps.lp")
     prob.solve()
-    print("Status:", LpStatus[prob.status])
 
-    list_vars = ['RockProbability', 'PaperProbability', 'ScissorProbability']
-    result = dict()
-    for var in list_vars:
-        result[var] = 0
+    if verbose:
+        print("Status:", LpStatus[prob.status])
+        for v in prob.variables():
+            print(v.name, "=", v.varValue)
 
-    for v in prob.variables():
-        print(v.name, "=", v.varValue)
-        if v.name in list_vars:
-            result[v.name] = v.varValue
-
-    return list(result.values())
+    return [p_r.varValue, p_p.varValue, p_s.varValue], v.varValue
 
 
 if __name__ == '__main__':
     matrix = np.array([[0.0, 1.04, -2.07], [-1.04, 0.0, 1.57], [2.07, -1.57, 0.0]])
-    result = solve_rps(matrix)
-    print(result)
+    result, total_value = solve_rps(matrix)
+    print(f'\nMixed strategy: {result}\nTotal value: {total_value}')
+

### 2019-07-12 project_3/main.py
Adding train method, still TODO choose_actions

@@ -1,4 +1,5 @@
 import numpy as np
+from pulp import *
 import environment
 from collections import defaultdict
 
@@ -9,9 +10,56 @@ class MultiQ(object):
         self.selection = selection
         self.alpha = alpha
         self.gamma = gamma
+        self.Q_matrix = defaultdict(lambda: np.zeros((env.num_actions,
+                                                      env.num_actions)))
 
-    def train(self):
-        pass
+    def minimax(self, r):
+        prob = LpProblem("Foe-Q", LpMaximize)
+
+        v = LpVariable("TotalValue")
+        p_0 = LpVariable("Action_0_Probability_Put", 0, 1)
+        p_1 = LpVariable("Action_1_Probability_North", 0, 1)
+        p_2 = LpVariable("Action_2_Probability_East", 0, 1)
+        p_3 = LpVariable("Action_3_Probability_South", 0, 1)
+        p_4 = LpVariable("Action_4_Probability_West", 0, 1)
+
+        prob += v
+        prob += p_0 + p_1 + p_2 + p_3 + p_4 == 1
+        for c in range(r.shape[1]):
+            prob += r[0][c] * p_0 + r[1][c] * p_1 + r[2][c] * p_2 + r[3][c] * \
+                    p_3 + r[4][c] * p_4 >= v
+
+        prob.solve()
+        return [p_0.varValue, p_1.varValue, p_2.varValue, p_3.varValue,
+                p_4.varValue], v.varValue
+
+    def choose_actions(self, epsilon, prob_actions=None):
+        # TODO select actions following ε-greedy policy
+        a0 = np.random.randint(self.env.num_actions)
+        a1 = np.random.randint(self.env.num_actions)
+        return a0, a1
+
+    def train(self, num_episodes=1000, render=False, epsilon_decay=0.99):
+        epsilon = 1.0
+        for i_episode in range(num_episodes):
+            state = self.env.reset()
+            done = False
+            a0, a1 = self.choose_actions(state, epsilon)
+            while not done:
+                if render:
+                    self.env.render()
+
+                next_state, rewards, done, info = self.env.step(a0, a1)
+
+                prob_actions, v = self.minimax(self.Q_matrix[next_state])
+                self.Q_matrix[state][a0][a1] = (1 - self.alpha) * self.Q_matrix[state][a0][a1] + self.alpha * ((1 - self.gamma) * rewards[0] + self.gamma * v)
+
+                state = next_state
+                epsilon *= epsilon_decay
[…]

### 2019-07-21 project_3/agents.py
Bug fixes in uCEQ

@@ -1,0 +1,370 @@
+import numpy as np
+import pandas as pd
+from pulp import *
+import environment
+from collections import defaultdict
+from abc import ABC, abstractmethod
+
+
+class BaseLearner(ABC):
+    def __init__(self, alpha0=0.1, alpha_min=0.001, alpha_decay=0.999995,
+                 gamma=0.9, state_s=((0, 1), (0, 2), 0)):
+        self.env = environment.Soccer()
+        self.alpha0 = alpha0
+        self.alpha_min = alpha_min
+        self.alpha_decay = alpha_decay
+        self.gamma = gamma
+        self.state_s = state_s
+
+    @abstractmethod
+    def choose_actions(self, epsilon, state):
+        pass
+
+    @abstractmethod
+    def update_Q_Pi(self, state, a0, a1, next_state, rewards, done, alpha):
+        pass
+
+    def train(self, num_timesteps=1e6, epsilon_decay=0.999995,
+              min_epsilon=0.001, print_same_line=True, log_filename='log.csv'):
+        state_s_visits = 0
+        error = 0.
+        count_errors = 0
+        epsilon = 1.0
+        t = 0
+        i_episode = 0
+        wins = [0, 0]
+        alpha = self.alpha0
+
+        df = pd.DataFrame(columns=['timestep', 'epsilon', 'alpha', 'error'])
+
+        while t <= num_timesteps:
+            i_episode += 1
+            state = self.env.reset()
+            while True:
+                a0, a1 = self.choose_actions(epsilon, state)
+                if (state == self.state_s) and (a1 == 3):
+                    state_s_visits += 1
+
+                next_state, rewards, done, info = self.env.step(a0, a1)
+
+                old_Q1, new_Q1 = self.update_Q_Pi(state, a0, a1, next_state,
+                                                  rewards, done, alpha)
+
+                if new_Q1 != old_Q1:
+                    error = abs(new_Q1 - old_Q1)
+                    df.loc[count_errors] = [int(t), epsilon, alpha, error]
+                    df.to_csv(log_filename, index=False)
+                    count_errors += 1
+
+                # Compute statistics
+                if rewards[0] == 100:
+                    wins[0] += 1
+                elif rewards[1] == 100:
+                    wins[1] += 1
+
+                # Log data in-screen
+                if t > 0 and print_same_line:
+                    sys.stdout.write(b'\033[3A'.decode())
+
[…]

### 2019-07-21 project_3/playground.py

@@ -1,5 +1,8 @@
 import numpy as np
 from collections import defaultdict
+import random
+from cvxopt import matrix, solvers
+import pulp
 
 
 def choice_experiments():
@@ -19,9 +22,81 @@ def remove_negatives():
     print(y)
 
 
-if __name__ == '__main__':
+def argmax():
     Q1 = defaultdict(lambda: np.zeros(5))
     print(Q1[0])
     Q1[0] = np.random.rand(5)
     print(Q1[0])
     print(np.argmax(Q1[0]))
+
+
+def adding_rows():
+    np.random.seed(1)
+    random.seed(1)
+    x = np.random.randint(4, size=(5, 5))
+    print(x)
+    print(np.sum(x, axis=1))
+
+
+def loop_i_j():
+    for i in range(5):
+        for j in range(5):
+            if (i == j) or (j < i):
+                continue
+
+            print(f'{i}-{j}')
+
+
+def experiment():
+    np.random.seed(1)
+    random.seed(1)
+    x = np.random.randint(4, size=(5, 5))
+    print(x)
+    probs = np.random.rand(5)
+    print(probs)
+    print(probs * x)
+    print(np.sum(probs * x, axis=1))
+    print(np.max(np.sum(probs * x, axis=1)))
+
+
+def lp1():
+    c = matrix([-1., -1.])
+    G = matrix([[4., 2., -5.], [-1., 1., 2.]])
+    h = matrix([8., 10., 2.])
+    solution = solvers.lp(c, G, h)
+    print(solution['x'])
+
+
+if __name__ == '__main__':
+    c = matrix([1., 1., 1., 1.])
+    G = matrix([[1., 0., 1., 0.], [-2., 0., 0., -1.],
+                [0., -1., -2., 0.], [0., 2., 0., 2.]])
+    h = matrix([0., 0., 0., 0.])
+    A = matrix([[1.], [1.], [1.], [1.]])
+    b = matrix(1.)
+    solution = solvers.lp(c, G, h, A, b, solver='glpk')
+    print(solution['x'])
+
+    # Pulp
+    prob = pulp.LpProblem("uCE-Q", pulp.LpMaximize)
+    p_1 = pulp.LpVariable("Action_0_Probability_Put", 0, 1, pulp.LpContinuous)
[…]

### 2019-07-21 project_3/playground2.py

@@ -1,0 +1,34 @@
+from cvxopt import matrix, solvers
+
+
+def solve_chicken(solver=None):
+    c = matrix([1., 1., 1., 1.])
+    G = matrix([[1., 0., 1., 0.], [-2., 0., 0., -1.], [0., -1., -2., 0.],
+                [0., 2., 0., 2.]])
+    h = matrix([0., 0., 0., 0.])
+
+    A = matrix([[1.], [1.], [1.], [1.]])
+    b = matrix(1.)
+
+    return solvers.lp(c, G, h, A, b, solver=solver)
+
+
+def solve_chicken2(solver=None):
+    c = matrix([1., 1., 1., 1.])
+    G = matrix([[1., 1., 0., 0., -1.], [-2., 0., 0., -1., -1.],
+                [0., -2., -1., 0., -1.], [0., 0., -2., -2., -1.]])
+    h = matrix([0., 0., 0., 0., -1.])
+
+    return solvers.lp(c, G, h, solver=solver)
+
+
+if __name__ == '__main__':
+    # suppress output
+    solvers.options['show_progress'] = False
+    solvers.options['glpk'] = {'LPX_K_MSGLEV': 0, 'msg_lev': "GLP_MSG_OFF"}
+
+    print("objective = {}".format(solve_chicken('glpk')['primal objective']))
+    print("x = {}".format(solve_chicken('glpk')['x']))
+
+    print("objective = {}".format(solve_chicken2('glpk')['primal objective']))
+    print("x = {}".format(solve_chicken2('glpk')['x']))

### 2019-07-21 project_3/playground3.py

@@ -1,0 +1,94 @@
+import numpy as np
+from cvxopt import matrix, solvers
+import pulp
+
+
+def ce(A, solver=None):
+    num_vars = len(A)
+    # maximize matrix c
+    c = [sum(i) for i in A] # sum of payoffs for both players
+    c = np.array(c, dtype="float")
+    c = matrix(c)
+    c *= -1 # cvxopt minimizes so *-1 to maximize
+    # constraints G*x <= h
+    G = build_ce_constraints(A=A)
+    print(G)
+    G = np.vstack([G, np.eye(num_vars) * -1]) # > 0 constraint for all vars
+    h_size = len(G)
+    G = matrix(G)
+    h = [0 for i in range(h_size)]
+    h = np.array(h, dtype="float")
+    h = matrix(h)
+    # contraints Ax = b
+    A = [1 for i in range(num_vars)]
+    A = np.matrix(A, dtype="float")
+    A = matrix(A)
+    b = np.matrix(1, dtype="float")
+    b = matrix(b)
+    sol = solvers.lp(c=c, G=G, h=h, A=A, b=b, solver=solver)
+    return sol
+
+
+def build_ce_constraints(A):
+    num_vars = int(len(A) ** (1/2))
+    G = []
+    # row player
+    for i in range(num_vars): # action row i
+        for j in range(num_vars): # action row j
+            if i != j:
+                constraints = [0 for i in A]
+                base_idx = i * num_vars
+                comp_idx = j * num_vars
+                for k in range(num_vars):
+                    constraints[base_idx+k] = (- A[base_idx+k][0]
+                                               + A[comp_idx+k][0])
+                G += [constraints]
+    # col player
+    for i in range(num_vars): # action column i
+        for j in range(num_vars): # action column j
+            if i != j:
+                constraints = [0 for i in A]
+                for k in range(num_vars):
+                    constraints[i + (k * num_vars)] = (
+                            - A[i + (k * num_vars)][1]
+                            + A[j + (k * num_vars)][1])
+                G += [constraints]
+    return np.matrix(G, dtype="float")
+
+
+def pulp_ce(A):
+    vars = pulp.LpVariable.dicts("Util", range(4), 0, 1.0)
+
+    # Create the 'prob' variable to contain the problem data
[…]

### 2019-07-21 project_3/playground4.py

@@ -1,0 +1,79 @@
+from cvxopt import matrix, solvers
+from cvxopt.base import spmatrix, spdiag
+from cvxopt.modeling import dot, op, variable
+import numpy as np
+
+solvers.options['show_progress'] = False
+solvers.options['msg_lev'] = 'GLP_MSG_OFF'
+solvers.options['show_progress'] = False  # disable solver output
+solvers.options['glpk'] = {'msg_lev': 'GLP_MSG_OFF'}  # cvxopt 1.1.8
+solvers.options['LPX_K_MSGLEV'] = 0  # previous versions
+
+
+# https://github.com/adam5ny/blogs/blob/master/cvxopt/cvxopt_examples.py
+def ce(Q, opQ, solver=None):
+    na = Q.shape[0]
+    nvars = na ** 2
+
+    Q_flat = Q.flatten()
+    opQ_flat = opQ.flatten()
+
+    # Minimize matrix c (*=-1 to maximize)
+    c = -np.array(Q_flat + opQ_flat, dtype="float")
+    c = matrix(c)
+
+    # Inequality constraints G*x <= h
+    G = np.empty((0, nvars))
+
+    # Player constraints
+    for i in range(na):  # action row i
+        for j in range(na):  # action row j
+            if i == j: continue
+            constraint = np.zeros(nvars)
+            base_idx = i * na
+            comp_idx = j * na
+            for _ in range(na):
+                constraint[base_idx + _] = Q_flat[comp_idx + _] - Q_flat[base_idx + _]
+            G = np.vstack([G, constraint])
+
+    # Opponent constraints
+    Gopp = np.empty((0, nvars))
+    for i in range(na):  # action row i
+        for j in range(na):  # action row j
+            if i == j: continue
+            constraint = np.zeros(nvars)
+            for _ in range(na):
+                # constraint[base_idx + j * _] = opQ_flat[comp_idx + _] - opQ_flat[base_idx + _]
+                constraint[i + _ * na] = opQ_flat[j + (_ * na)] - opQ_flat[i + (_ * na)]
+            Gopp = np.vstack([Gopp, constraint])
+
+    G = np.vstack([G, Gopp])
+    G = np.matrix(G, dtype="float")
+    G = np.vstack([G, -1. * np.eye(nvars)])
+    h_size = len(G)
+    G = matrix(G)
[…]

### 2019-07-21 project_3/uceq.py

@@ -1,0 +1,252 @@
+import numpy as np
+import pandas as pd
+from cvxopt import matrix, solvers
+import environment
+import sys
+from collections import defaultdict
+import pulp
+
+
+solvers.options['show_progress'] = False
+solvers.options['msg_lev'] = 'GLP_MSG_OFF'
+solvers.options['show_progress'] = False  # disable solver output
+solvers.options['glpk'] = {'msg_lev': 'GLP_MSG_OFF'}  # cvxopt 1.1.8
+solvers.options['LPX_K_MSGLEV'] = 0  # previous versions
+np.set_printoptions(precision=3)
+
+
+class uCEQ(object):
+    def __init__(self, alpha0=0.1, alpha_min=0.001, alpha_decay=0.999995,
+                 gamma=0.9):
+        self.env = environment.Soccer()
+        self.alpha0 = alpha0
+        self.alpha_min = alpha_min
+        self.alpha_decay = alpha_decay
+        self.gamma = gamma
+        self.Q0 = defaultdict(lambda: np.zeros((self.env.num_actions,
+                                                self.env.num_actions)))
+        self.Q1 = defaultdict(lambda: np.zeros((self.env.num_actions,
+                                                self.env.num_actions)))
+        self.P0 = defaultdict(lambda: np.ones(self.env.num_actions) /
+                                      self.env.num_actions)
+        self.P1 = defaultdict(lambda: np.ones(self.env.num_actions) /
+                                      self.env.num_actions)
+
+    def pulp_ce(self, r0, r1):
+        num_vars = self.env.num_actions ** 2
+        vars = pulp.LpVariable.dicts("Util", range(num_vars), 0, 1.0)
+
+        # Create the 'prob' variable to contain the problem data
+        prob = pulp.LpProblem("Chicken problem", pulp.LpMaximize)
+        sum_rewards = r0.flatten() + r1.flatten()
+
+        prob += pulp.lpSum([vars[i] * sum_rewards[i] for i in range(num_vars)])
+        prob += pulp.lpSum([vars[i] for i in range(num_vars)]) == 1
+
+        G = self.build_ce_constraints(r0, r1)
+        for r in range(G.shape[0]):
+            prob += pulp.lpSum([G[r, c] * vars[c] for c in range(G.shape[1])]) <= 0
+
+        prob.solve()
+        probabilities = np.array([vars[i].varValue for i in range(num_vars)])
+        probabilities -= probabilities.min() + 0.
[…]

### 2019-07-22 project_3/agents.py
Adding report

@@ -1,9 +1,27 @@
+# -*- coding: utf-8 -*-
+"""OMSCS Reinforcement Learning - CS-7642-O03 - Project #3 Solution
+
+This code solves Project #3 from OMSCS Reinforcement Learning - CS-7642-O03,
+generating the data used in the report. Running it is straightforward:
+
+    $ python agents.py {qlearner, friendq, foeq, or uceq}
+
+This will train each of the agents, generating the CSV data used to create the
+the figure with all 4 charts in the file 'images/figure.png', used in the
+report.
+
+Created by Carlos Souza (souza@gatech.edu)
+Jul-2019
+
+"""
+
 import numpy as np
 import pandas as pd
 from pulp import *
 import environment
 from collections import defaultdict
 from abc import ABC, abstractmethod
+import argparse
 
 
 class BaseLearner(ABC):
@@ -299,7 +317,7 @@ class uCEQ(BaseLearner):
                                               self.env.num_actions) / \
             probabilities.sum(0)
         pi = np.sum(probabilities, axis=1)
-        v = np.sum(probabilities * r0)
+        v = np.max(np.sum(probabilities * r0, axis=1))
         return pi, v
 
     def build_ce_constraints(self, r0, r1):
@@ -368,3 +386,27 @@ class uCEQ(BaseLearner):
 
         new_Q1 = self.Q1[self.state_s][3][0]
         return old_Q1, new_Q1
+
+
+if __name__ == '__main__':
+    parser = argparse.ArgumentParser(description='Generate Project 3 charts.')
+    parser.add_argument('agent', choices=['qlearner', 'friendq', 'foeq',
+                                          'uceq'])
+    parser.add_argument('-f', '--file', default=None)
+    args = parser.parse_args()
+
+    solver, filename = None, None
+    if args.agent == 'qlearner':
+        solver = QLearner()
+        filename = args.file if args.file is not None else 'qlearner.csv'
+    elif args.agent == 'friendq':
+        solver = FriendQ()
+        filename = args.file if args.file is not None else 'friendq.csv'
+    elif args.agent == 'foeq':
+        solver = FoeQ()
+        filename = args.file if args.file is not None else 'foeq.csv'
+    elif args.agent == 'uceq':
+        solver = uCEQ()
+        filename = args.file if args.file is not None else 'uceq.csv'
+
+    solver.train(log_filename=filename)

### 2019-07-22 project_3/environment.py

@@ -4,6 +4,9 @@
 This module implements Soccer environment as described in Greenwald & Hall's
 `Correlated-Q Learning paper`_. Implementation follows `OpenAI Gym standards`_.
 
+Created by Carlos Souza (souza@gatech.edu)
+Jul-2019
+
 .. _Correlated-Q Learning paper:
    https://www.aaai.org/Papers/ICML/2003/ICML03-034.pdf
 

### 2019-07-22 project_3/uceq.py

@@ -200,11 +200,11 @@ class uCEQ(object):
                     self.Q1[state][a1][a0] = (1 - alpha) * self.Q1[state][a1][a0] + alpha * rewards[1]
                 else:
                     prob_actions0, v0 = self.pulp_ce(self.Q0[next_state], self.Q1[next_state])
-                    v0 = np.max(self.Q0[next_state])
+                    #v0 = np.max(self.Q0[next_state])
                     self.P0[next_state] = prob_actions0
 
                     prob_actions1, v1 = self.pulp_ce(self.Q1[next_state], self.Q0[next_state])
-                    v1 = np.max(self.Q1[next_state])
+                    #v1 = np.max(self.Q1[next_state])
                     self.P1[next_state] = prob_actions1
 
                     self.Q0[state][a0][a1] = (1 - alpha) * self.Q0[state][a0][a1] + alpha * (rewards[0] + self.gamma * v0)

### 2019-07-22 project_3/agents.py
Adding final paper and project

@@ -7,8 +7,7 @@ generating the data used in the report. Running it is straightforward:
     $ python agents.py {qlearner, friendq, foeq, or uceq}
 
 This will train each of the agents, generating the CSV data used to create the
-the figure with all 4 charts in the file 'images/figure.png', used in the
-report.
+figure with all 4 charts in the file 'images/figure2.png', used in the report.
 
 Created by Carlos Souza (souza@gatech.edu)
 Jul-2019
